---
title: "bspline_tensor"
author: "Ryan Peterson"
date: "2024-08-20"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library}
suppressMessages(library(LatticeKrig))

suppressMessages(library(fda)) #for splineDesign
```


TODO: update for actual application in LatticeKrig

Goal is to create a surface from b-spline tensor products across a lattice.

Steps:

1. Define/create cardinal b-spline, and apply tensor product.
  a. Using cubicSplineBasis function define on 1-d grid
  b. create tensor product for a single `basis' function.

  
  
2. Create lattice then center tensor basis on each knot.
  a. evaluate and visualize a 2-d grid.
  
3. Add multiple smaller layers from LatticeKrig
 

Ancillary/Older work:

Make adjustments to `naturalSplineBasis` function to create interior cardinal b-splines


```{r functions}
setwd("~/COS_LK")
source("DF_LK/R/naturalSplineBasis.R") #fix this for our specific purpose

source("DF_LK/R/bspline_poly.R")
```


# 1 

```{r 1d grid}

base_knots <- c(-3:3) #add in buffer for no reason (not needed, since data is centered)
x_grid_base <- seq(-2.5, 2.5, 0.01)
cubicSplineBasis(x_grid_base)


```

Showing partition of unity example.

```{r partition_example}
x_grid <- seq( 0.99,6,.01)

min_knot <- floor(min(x_grid))
max_knot <- ceiling(max(x_grid))
buffer <- c(min_knot - 1, max_knot + 1) 

#center of each internal knot
base_knots <- seq(min_knot, max_knot, 1)

#center of each bspline function
x_knots <- seq(min(buffer), max(buffer), 1) #including buffer knots


test_bsplines <- matrix(NA, nrow = length(x_grid))
for (j in x_knots) {
  test_bsplines <- cbind(test_bsplines, cubicSplineBasis(x_grid - j))
}
test_bsplines <- test_bsplines[,-1]

#test plot, gray line shows partition of unity
plot(range(x_grid), c(0,1), type = "n", xlab = "x", ylab = "")
abline(v = base_knots, lty = 2)
matlines(x_grid, test_bsplines, ylim = c(0,1), lty = 1)
lines(x_grid, rowSums(test_bsplines), col = "gray", lwd = 2)
```


# Older Work

```{r raw_function}
naturalSplineBasis <- function(sGrid,
                               sKnots,
                               degree = 3,
                               derivative = 0) {
  boundaryKnots<- c( min(sKnots),max(sKnots))
  sKnots0<- c( rep( boundaryKnots[1],degree),sort(sKnots),
               rep( boundaryKnots[2],degree) )
  testRight<- sGrid < min(sKnots) 
  testLeft <- sGrid > max(sKnots)             
  if( any(testRight |testLeft) )
  {stop("some points for evaluation outside knot range.")}
               
  basis <- splineDesign(sKnots0, sGrid,
                        ord= degree+1, outer.ok=TRUE,
                        derivs=derivative)
  # set up constraints to enforce natural BCs.
  #require library(fda)
  const <- splineDesign(sKnots0, boundaryKnots, ord = degree+1,
                        derivs = c(2,2)) 
  qr.const <- qr(t(const))
  QBasis<- t(qr.qty( qr.const, t(basis) ))
  basis <- QBasis[,-(1:2)]
  basis
  
  return( basis )
  
}
```

# 1. Cardinal B-spline

Degree 3 cardinal B-spline (order 4)



```{r}
x2knots <- seq(-4, 4, 1)
#x2Grid <- seq(-1,1,0.01)
x2Grid <- seq( -4, 4, .01)
look <- naturalSplineBasis(x2Grid, x2knots, degree=3) #order 3, degree 2
c_1 <- 3/2 #scaling coeff
look1 <- c_1 * look[,4] #coefficient should work here too

plot(x2Grid, look1, type="l")
abline(v = seq(-4, 4, 1), lty = 2)
```

## . Tensor Partition of Unity

Test that the 2-D version of our tensor b-splines will follow our expected partition of unity.

- Centered on every knot on the lattice with integer spacing

- This will require a minimum of two buffer knots on all boundaries.

```{r}
look2 <- look[,3]
#TODO: fix this, something is not quite working correctly, and if it is, I need to shift and scale it.
tSpline1 <- outer(look1, look1, "*")
tSpline2 <- outer(look2, look2, "*")

tSpline3 <- outer(look1, look2, "*")
tSpline4 <- outer(look2, look1, "*")

test_surface <- tSpline1 + tSpline2 

max(look1)^2
max(tSpline1)

imagePlot(test_surface)


test_grid <- expand.grid(X = x2knots, Y = x2knots)

plot(test_grid$X, as.numeric(test_grid$Y), 
     xlab = "X", ylab = "Y",
     pch = 19, xlim = c(-3, 3), ylim = c(-3, 3))
rect(-2, -2, 2, 2, border = "darkmagenta", lwd = 3)
for (i in -4:4) {
  abline(h= i, lty = 2, lwd = 0.5, col = "darkgrey")
  abline(v= i, lty = 2, lwd = 0.5, col = "darkgrey")
}


```

