---
title: "COS_Base"
author: "Ryan Peterson"
date: "2023-08-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
suppressMessages(library(LatticeKrig))
```


Goal: begin work on Change of Support problem/code for LatticeKrig

Consider estimates from Lattice Krig where,
$$\textbf{z} = U \textbf{d} + X \textbf{c} + \textbf{e}.$$

Thus, we want to create a $U$ and a $X$ matrix for LatticeKrig from out block data.

```{r test_data}
load("COSExampleData.rda")
```

Test data is generated from :
`createExampleCOSLatticeKrig1.R`
Which: creates the synthetic COS data sets to test

To change the underlying surface that the data is generated from adjust the function:
`FUNTrue` in `createExampleCOSLatticeKrig1.R`

This is the true function of the surface we are trying to find from the block data.


Loaded test data includes:

-polyGroups
 polyGroups is a list with each component being a 2 column matrix with the outlines of the region. 
 5 rows and 2 columns that create an outline of a region (block) 
 these 5 points draw the outline of our object. (beginning and end point are included even though they are the same)

-polyGroupsRandom
  
--Primary randomly generated blocks/areal units

-polyGroupsRandomBig

Initial Exploration:

-To look at data fusion we want either split the polyGroups or create centroids from each polyGroup
--I think the centroids have been found in a file somewhere.

# Regular Tiles

From `COSExampleRegularTiles.R` script

Setup for this script includes several function files

```{r function_files}
source("COSExample/R/basisIntegral.R")
source("COSExample/R/integralPolyFunction.R")
# patch in newer version of LK functions
source("COSExample/R/LatticeKrig.R")
source("COSExample/R/LKrigFindLambda.R")
source("COSExample/R/print.LatticeKrig.R" )
```

We begin by creating a Setup file for LatticeKrig, with only a single layer, and an a weight only  slightly above minimum (a.wght > 4). This will also create a $25 \times 25$ grid of values on our given `sDomain`.

`sDomain` is the spatial domain of interest where we have a easting (x axis) from $-1$ to $1$ and a northing (y axis) of $0$ to $2$. 

```{r}
#uses lattice krig setup file
##test later wih a NC.buffer
LKinfo<- LKrigSetup(sDomain, NC=25, nlevel=1, a.wght = 4.1,
                    NC.buffer = 0, normalize = FALSE)

#creates a couple functions, not sure what they do yet.
FUNX<- function(s){
  s[,1]
}

FUNY<- function(s){
  s[,2]
}
```


Review what this does again:
Creates matrix of fixed effect integrals, where we look at the integralPolyFunction
See the `integralPolyFunction.R` for more details

Notes from `integralPolyFunction.R`:
  # finds approx integral by keeping basis function fixed
  # and translating and scaling the region.
  # integrals returned in spind sparse matrix format
  
My dumb idea (MDI):
- Include monte carlo integration if it isn't already included.



Fixed effect matrix $U$, where $U_1$ is a default constant function, $U_2$ is for x values, and $U_3$ is for y values. Thus our $U$ matrix has some constant value, with x and y values.

(Note: the parameter function function(s){s[,1]}, is a function that strips the first column of data that is passed to it. For example if $s$ is in the form $(x,y)$, then function(s){s[,1]} brings in the column of $x$s and function(s){s[,2]} brings in the column of $y$s.)

```{r}
# create matrix for fixed effects integrals
# (a linear function of coordinates in this case)
U1<- integralPolyFunction(polyGroups, M = 200)  # default is constant function 
U2<- integralPolyFunction(polyGroups,
      FUN=function(s){s[,1]}, M = 200)
U3<- integralPolyFunction(polyGroups,
      FUN=function(s){s[,2]}, M = 200)
U<- cbind( U1,U2,U3) # U can be a dense matrix for LK
```


Integrals of basis function for the $X$ matrix:

`basisIntegral` produces sparse index matrix for the `spind2spam` function. Thus we have a sparse matrix $X$ from? Answer: this is from the random effects model, so that we are using the basis integral for the basis functions. And thus we get a sparse matrix from this data since this is the primary goal of LatticeKrig.

```{r}
# integrals of basis functions
X<- basisIntegral( polyGroups, LKinfo, M = 400)
X<- spind2spam(X)
```


From the last two sections we see that I am missing some understanding of how `integralPolyFunction.R` and `basisIntegral.R` work, so we have the file: `COS_Integrals.Rmd` to work through this, and Line integrals.

Fitting the model with LatticeKrig, where we have a $U$ and $X$ generated above using `integralPolyFunction.R` and `basisIntegral.R`.

```{r}
# fit model using fixed lambda
fit0<- LKrig( sDomain,yObs, U=U, X=X, LKinfo=LKinfo,
              lambda=.01)

# estimating lambda
fit1<- LatticeKrig( sDomain,yObs, U=U, X=X, LKinfo=LKinfo)
                    
                   

# estimating lambda and the a.wght parameter

fit2<- LatticeKrig( sDomain,yObs, U=U, X=X, LKinfo=LKinfo,
                    findAwght=TRUE)
                   

pdf("COSExampleFittedSurfaces.pdf", width=6, height=6)
set.panel(2,2)
zlim<- range( gTrue)
surface(as.surface( gridTrue, gTrue) , 
        zlim =zlim, col=viridis(256))
title("true surface")
surface( fit0, zlim =zlim)
title("lambda=.01  a.wght=4.1")
surface( fit1, zlim =zlim)
title("lambda  MLE  a.wght=4.1")
surface( fit2, zlim =zlim)
title("lambda and a.wght MLEs")
```

## Random Big Prediction

Create a prediction from the random big polygons

```{r random_big}

```

