---
title: "COS_Integrals"
author: "Ryan Peterson"
date: "2023-10-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
suppressMessages(library(LatticeKrig))
```


Goal: to understand how `integralPolyFunction.R` and `basisIntegral.R` work. With extensions to using line integrals.

When done we can convert this to an R script that has all the integral functions.

Stretch goals:

-pass in shapefiles as polyGroups



```{r}
#importing polyGroup data to work with.
load("COSExampleData.rda")
```

LKinfo: passed from latticeKrig, set up LKinfo first

```{r LKinfo}
#setup for LKinfo (which should be used in the function)
LKinfo<- LKrigSetup(sDomain, NC=25, nlevel=1, a.wght = 4.1,
                    NC.buffer = 0, normalize = FALSE)
```


# integralPolyFunction
(integrals for fixed effects)
Function parameters:
polyGroups: polygon regions/geometric objects in the form of 2 column matrix that creates a closed curve; 
FUN: some function, default: NULL; 
-first parameter of the function must take in grid data (location data in the form (x,y))
M: number of points used to integrate over, default = 200


```{r functions}
#creates a couple functions, not sure what they do yet.
FUNX<- function(s){
  s[,1]
}

FUNY<- function(s){
  s[,2]
}
```


```{r integralPolyFunction}
#build without the function wrapper

##from `integralPolyFunction.R`:
# "polyGroups is a list with each component being a 2 column matrix with the outlines of the region."
# "uses the basis information in a LKinfo LKinfo"
# "M is the number of points used to integrate over basis function"

#pass parameters (for default constant function, update for x and y later):
polyGroups <- polyGroups
FUN <- NULL
M <- 200



#length of polyGroups matrix (number of geometric objects)
N1 <- length(polyGroups)

#create empty object the length of the total number of objects
theIntegrals<-rep(NA, N1)

#perform below loop on a samller set of values

# loop over the polygon regions
for (j in 1:N1) {
  #select a geom object
  polyTmp <- (polyGroups)[[j]]
  
  # creates list of M points between the min x (left) and max x (right),
  # M points between min y (bottom) and max y (top)
  ##appears to only hand rectangular objects, however, this might also work for other objects.
  polyGridList <- list(
                   x= seq(min(polyTmp[,1]), max(polyTmp[,1]), length.out=M),
                   y= seq(min(polyTmp[,2]), max(polyTmp[,2]), length.out=M)
                   )
  print(polyGridList)
  
  #distance between each integration point (M =200)
  dxP <- polyGridList$x[2] - polyGridList$x[1]
  dyP <-polyGridList$y[2] - polyGridList$y[1]
  
  #surface of M^2 points and 2 * M^2 elements
  polyGrid <- make.surface.grid( polyGridList)
  
  #fields function to determine if the points of polyGrid are inside the polygon polyTmp
  ind <- in.poly(polyGrid, polyTmp)
  polyGrid<- polyGrid[ind,]
}
```


```{r test_block}
M <- 200
j <- 2

polyTmp <- (polyGroups)[[j]]
polyGridList <- list(
                x= seq(min(polyTmp[,1]), max(polyTmp[,1]), length.out=M),
                y= seq(min(polyTmp[,2]), max(polyTmp[,2]), length.out=M)
                )
dxP<-  polyGridList$x[2] - polyGridList$x[1]
dyP<-  polyGridList$y[2] - polyGridList$y[1]


#surface grid of 40,000 (200 x 200) points (80,000 elements (e.g. x,y for each point))
polyGrid <- make.surface.grid( polyGridList)

#TODO: try to output this visually with the polygon
#look up USGS work for a similar example
ind <- in.poly(polyGrid, polyTmp)

#selects for only values (rows) of the grid that are inside the polygon
polyGrid <- polyGrid[ind,]

#if else conditionals
#if !is.null(FUN), if that function is not null, then
look <- FUN(polyGrid, ... ) #passing grid data to the function
the_integral <- sum( look)*dxP*dyP 

#else (that is FUN = NULL) (unit blocks/points)
the_integral <- sum( ind)*dxP*dyP
```

# basisIntegral
(integral for random effects, from basis functions and GMRF)
Function parameters:


```{r basisIntegral}
#parameters
polyGroups <- polyGroups
LKinfo <- LKinfo
M = 200
check=TRUE

# setup basis function info

#number of levels for the basis functions (from LKInfo set up)   
nLevel <- LKinfo$nlevel

  
```

