---
title: "COS_synthetic"
author: "Ryan Peterson"
date: "2023-11-10"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
suppressMessages(library( fields))
suppressMessages(library( LatticeKrig))
```

Goal: create various forms of our synthetic data to test with

```{r functions}
source( "basisIntegral.R")
source("integralPolyFunction.R")
```

Visualizations for Random Big 1 & 2, Grid, and True Surface are all in this Document.

TODO:

1. Change the true surface (DONE)

2. Create variations in the sampling and increase the sample size to 5000 (DONE)

3. Get visualizations for true surface and sampling (In Progress)

4. Get are of each sample block

Create polygons

```{r}
M1<- 32
M2<- 32
sDomain<- rbind( c( -1,-1),
                 c( 1,2))
tileGridX<- seq( sDomain[1,1],sDomain[2,1], length.out = M1+1)
tileGridY<- seq( sDomain[1,2],sDomain[2,2], length.out = M2+1)
tmpGrid<- list( x= tileGridX[1:M1] + diff(tileGridX)/2, 
                y= tileGridY[1:M2] + diff(tileGridY)/2
                )

# for plotting later on 
tileCenters<-  make.surface.grid( tmpGrid)
```

Polygon groups (different orientations of polygons)

```{r base_polygon}
polyGroups<- NULL
for( j in 1:M1){
  for ( k in 1:M2){
    newPoly<-  rbind( c(tileGridX[j], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k])
    )
    polyGroups<- c( polyGroups,list(newPoly))
  }
}

rm(j,k, newPoly)
```


```{r random_polygonsBIG}
#create two random big sets of polygons
polyGroupsRandomBig <- NULL
dx <- tileGridX[2] - tileGridX[1]
dy <- tileGridY[2] - tileGridY[1]
M<- 5000
set.seed(155)
for( i in 1:M){
    X1<- runif(1, sDomain[1,1], sDomain[2,1]) 
    Y1<- runif(1, sDomain[1,2], sDomain[2,2]) 
    ranScale<- runif( 1,.25,.5)
    X2<- min(c( X1+(2*dx*ranScale),sDomain[2,1] ))
    Y2<- min(c( Y1+(2*dy*ranScale),sDomain[2,2] ))
    
    newPoly<-  rbind( c(X1, Y1),
                      c(X2, Y1),
                      c(X2, Y2),
                      c(X1, Y2),
                      c(X1, Y1)
    )
    polyGroupsRandomBig<- c( polyGroupsRandomBig,list(newPoly))
}

polyGroupsRandomBig_2 <- NULL
dx <- tileGridX[2] - tileGridX[1]
dy <- tileGridY[2] - tileGridY[1]
M<- 5000
set.seed(200)
for( j in 1:M){
    X1<- runif(1, sDomain[1,1], sDomain[2,1]) 
    Y1<- runif(1, sDomain[1,2], sDomain[2,2]) 
    ranScale<- runif( 1,.20,.55)
    X2<- min(c( X1+(2*dx*ranScale),sDomain[2,1] ))
    Y2<- min(c( Y1+(2*dy*ranScale),sDomain[2,2] ))
    
    newPoly<-  rbind( c(X1, Y1),
                      c(X2, Y1),
                      c(X2, Y2),
                      c(X1, Y2),
                      c(X1, Y1)
    )
    polyGroupsRandomBig_2<- c( polyGroupsRandomBig_2, list(newPoly))
}

rm(M, i, j, newPoly)
```

```{r test_vis}
yTrueRandomBig<-  integralPolyFunction(polyGroupsRandomBig,
                                    FUN=FUNTrue,
                                    M = 400)

set.seed( 114)
n<- length( yTrueRandomBig)
tau<- .1
yObsRandomBig<- yTrueRandomBig +  tau*rnorm(n )

N<- length(polyGroupsRandomBig)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObsRandomBig, cScale)
plot( sDomain, type="n", axes = FALSE,
      xlab = "x", ylab = "y", main = "Dataset 1")
axis(1)
axis(2, las =1)
for ( k in 1:N){
  polyTmp<- polyGroupsRandomBig[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}
#addLegend(layout, col=cScale, zlim= range(yTrueRandomBig ))
```

```{r polyGroupsRandomBig_2_vis}
yTrueRandomBig_2 <- integralPolyFunction(polyGroupsRandomBig_2,
                                    FUN=FUNTrue,
                                    M = 400)

set.seed( 158)
n <- length( yTrueRandomBig_2)
tau <- .1
yObsRandomBig_2 <- yTrueRandomBig_2 +  tau*rnorm(n )

N<- length(polyGroupsRandomBig_2)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObsRandomBig_2, cScale)
plot( sDomain, type="n",  axes = FALSE,
      xlab = "x", ylab = "y", main = "Dataset 2")
axis(1)
axis(2, las =1)
for ( k in 1:N){
  polyTmp<- polyGroupsRandomBig_2[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}
```


```{r grid_polygon_vis}
yTrue<-  integralPolyFunction(polyGroups,
                              FUN=FUNTrue,
                              M = 400)
set.seed( 111)
n<- length( yTrue)
tau<- .1
yObs<- yTrue +  tau*rnorm(n )

N<- length(polyGroups)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObs, cScale)
layout<-setupLegend()
plot( sDomain, type="n")
for ( k in 1:N){
  polyTmp<- polyGroups[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}
```

```{r larger_grid_size}
M1<- 24
M2<- 24
sDomain<- rbind( c( -1,-1),
                 c( 1,2))
tileGridX<- seq( sDomain[1,1],sDomain[2,1], length.out = M1+1)
tileGridY<- seq( sDomain[1,2],sDomain[2,2], length.out = M2+1)
tmpGrid<- list( x= tileGridX[1:M1] + diff(tileGridX)/2, 
                y= tileGridY[1:M2] + diff(tileGridY)/2
                )

polyGroups2<- NULL
for( j in 1:M1){
  for ( k in 1:M2){
    newPoly<-  rbind( c(tileGridX[j], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k])
    )
    polyGroups2<- c( polyGroups2,list(newPoly))
  }
}

yTrue2<-  integralPolyFunction(polyGroups2,
                              FUN=FUNTrue,
                              M = 400)
set.seed( 123)
n<- length( yTrue2)
tau<- .1
yObs2<- yTrue2 +  tau*rnorm(n )

N<- length(polyGroups2)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObs2, cScale)
plot( sDomain, type="n",  axes = FALSE,
      xlab = "x", ylab = "y", main = "Grid Dataset")
axis(1)
axis(2, las =1)
for ( k in 1:N){
  polyTmp<- polyGroups2[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}

save(polyGroups2, yTrue2, yObs2, 
     polyGroupsRandomBig_2, yTrueRandomBig_2, yObsRandomBig_2,
     polyGroupsRandomBig, yTrueRandomBig, yObsRandomBig,  file = "COSExampleNEW.rda")
```


```{r old}
polyGroups<- NULL
for( j in 1:M1){
  for ( k in 1:M2){
    newPoly<-  rbind( c(tileGridX[j], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k])
    )
    polyGroups<- c( polyGroups,list(newPoly))
  }
}

polyGroupsRandom<- NULL
set.seed(111)
dx<- tileGridX[2] - tileGridX[1]
dy<- tileGridY[2] - tileGridY[1]
M<- M1*M2
for( i in 1: (M)){
  X1<- runif(1, sDomain[1,1], sDomain[2,1]) 
  Y1<- runif(1, sDomain[1,2], sDomain[2,2]) 
  ranScale<- 1.0
  X2<- min(c( X1+dx*ranScale,sDomain[2,1] ))
  Y2<- min(c( Y1+dy*ranScale,sDomain[2,2] ))
  
  newPoly<-  rbind( c(X1, Y1),
                    c(X2, Y1),
                    c(X2, Y2),
                    c(X1, Y2),
                    c(X1, Y1)
  )
  
 polyGroupsRandom<- c( polyGroupsRandom,list(newPoly))
}
```




```{r random2}
polyGroupsRandom2 <- NULL
set.seed(555)
dx<- tileGridX[2] - tileGridX[1]
dy<- tileGridY[2] - tileGridY[1]
M<- M1*M2
for( i in 1: (M)){
  X1<- runif(1, sDomain[1,1], sDomain[2,1]) 
  Y1<- runif(1, sDomain[1,2], sDomain[2,2]) 
  ranScale<- 1.0
  X2<- min(c( X1+dx*ranScale,sDomain[2,1] ))
  Y2<- min(c( Y1+dy*ranScale,sDomain[2,2] ))
  
  newPoly<-  rbind( c(X1, Y1),
                    c(X2, Y1),
                    c(X2, Y2),
                    c(X1, Y2),
                    c(X1, Y1)
  )
  
 polyGroupsRandom2<- c( polyGroupsRandom2,list(newPoly))
}
```


Get area of each poly group

```{r area}
#assign an area to each polygon object (for polyGroups and polyGroupsRandom)

#using this for now, I'd like to not have this dependency
library(pracma)

n <- length(polyGroups)
area_matrix <- matrix(NA, nrow = n, ncol =1)
for (i in 1:n) {
  temp <- polyGroups[[i]]
  x1 <- temp[,1]
  y1 <- temp[,2]
  area_matrix[i] <- polyarea(x1,y1)
}

m <- length(polyGroupsRandom)
area_matrix_random <- matrix(NA, nrow = m, ncol =1)
for (i in 1:m) {
  temp <- polyGroupsRandom[[i]]
  x1 <- temp[,1]
  y1 <- temp[,2]
  area_matrix_random[i] <- polyarea(x1,y1)
}
```

True spatial field

```{r true_function}
#set up grid
bigGridList<- list( x= seq(-1,1,length.out=500), y= seq(0,2,length.out=500) )
dx<- bigGridList$x[2] - bigGridList$x[1]
dy<-bigGridList$y[2] - bigGridList$y[1]
bigGrid<- make.surface.grid( bigGridList)

# true function 
FUNTrue<- function(bigGrid){
D1<-  sqrt((bigGrid[,1] - 0.8)^2 + (bigGrid[,2] - 1.5)^2)
D2 <- sqrt((bigGrid[,1])^2 + (bigGrid[,2]-1)^2)
D3 <-  sqrt((bigGrid[,1] + 0.75)^2 + (bigGrid[,2] + 0.5)^2) 
g<-   100*(2*exp( -1*D1/0.5) +  1.75*exp( -1*D2/.5) + exp( -1*D3/0.2))
return(g)
}

yTrue <- integralPolyFunction(polyGroups,
                              FUN=FUNTrue,
                              M = 400)

gridTrue <- list( x= seq( sDomain[1,1],sDomain[2,1], length.out = 100),
                 y= seq( sDomain[1,2],sDomain[2,2], length.out = 100)
)
trueXY <- make.surface.grid( gridTrue )
gTrue <- FUNTrue(trueXY)


zlim<- range( gTrue)
surface(as.surface( gridTrue, gTrue) , 
        zlim =zlim, col=turbo(256))
title("True Surface")
```


```{r}
yTrue <- integralPolyFunction(polyGroups,
                              FUN=FUNTrue,
                              M = 400)
yTrueRandom <- integralPolyFunction(polyGroupsRandom,
                              FUN=FUNTrue,
                              M = 400)

yTrueRandom2 <- integralPolyFunction(polyGroupsRandom2,
                              FUN=FUNTrue,
                              M = 400)

gridTrue <- list( x= seq( sDomain[1,1],sDomain[2,1], length.out = 100),
                 y= seq( sDomain[1,2],sDomain[2,2], length.out = 100)
)
trueXY <- make.surface.grid( gridTrue )
gTrue <- FUNTrue(trueXY)

# make up the data
set.seed( 111)
n<- length( yTrue)
tau<- .1
yObs<- yTrue +  tau*rnorm(n )

set.seed( 112)
n<- length( yTrueRandom)
yObsRandom<- yTrueRandom +  tau*rnorm(n )

set.seed( 113)
n<- length( yTrueRandom2)
yObsRandom2<- yTrueRandom2 +  tau*rnorm(n )

#set.seed( 114)
#n<- length( yTrueRandomBig)
#yObsRandomBig<- yTrueRandomBig +  tau*rnorm(n )


```



```{r reg_wide}
N1 <- 8
N2 <- 8

tileGridX<- seq( sDomain[1,1],sDomain[2,1], length.out = N1+1)
tileGridY<- seq( sDomain[1,2],sDomain[2,2], length.out = N2+1)
tmpGrid<- list( x= tileGridX[1:N1] + diff(tileGridX)/2, 
                y= tileGridY[1:N2] + diff(tileGridY)/2
                )

polyGroupsWide<- NULL
for( j in 1:N1){
  for ( k in 1:N2){
    newPoly<-  rbind( c(tileGridX[j], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k]),
                      c(tileGridX[j+1], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k+1]),
                      c(tileGridX[j], tileGridY[k])
    )
    polyGroupsWide<- c( polyGroupsWide,list(newPoly))
  }
}

yTrueWide <- integralPolyFunction(polyGroupsWide,
                              FUN=FUNTrue,
                              M = 400)

# make up the data
set.seed( 111)
n<- length( yTrueWide)
tau<- .1
yObsWide<- yTrueWide +  tau*rnorm(n )

save(polyGroupsWide, yTrueWide, yObsWide, 
    polyGroupsRandom2, yTrueRandom2, yObsRandom2, file = "COSExampleData2.rda")

```


Add in some visualization to check

```{r}
N<- length(polyGroupsRandom)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObsRandom, cScale)
layout<-setupLegend()
plot( sDomain, type="n")
for ( k in 1:N){
  polyTmp<- polyGroupsRandom[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}


N<- length(polyGroupsRandom2)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObsRandom2, cScale)
layout<-setupLegend()
plot( sDomain, type="n")
for ( k in 1:N){
  polyTmp<- polyGroupsRandom2[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}

N<- length(polyGroupsRandom)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObsRandom, cScale)
layout<-setupLegend()
plot( sDomain, type="n")
for ( k in 1:N){
  polyTmp<- polyGroupsRandom[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}

```

