---
title: "COS_Base"
author: "Ryan Peterson"
date: "2023-08-22"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
suppressMessages(library(LatticeKrig))
```


Goal: begin work on Change of Support problem/code




```{r test_data}
load("COSExampleData.rda")
```

Loaded test data includes:

-polyGroups
 polyGroups is a list with each component being a 2 column matrix with the outlines of the region. 
 5 rows and 2 columns that create an outline of a region (block) 
 these 5 points draw the outline of our object. (beginning and end point are included even though they are the same)

-polyGroupsRandom

-polyGroupsRandomBig

Initial Exploration:

# Regular Tiles

From `COSExampleRegularTiles.R` script

Setup for this script includes several function files

```{r function_files}
source("R/basisIntegral.R")
source("R/integralPolyFunction.R")
# patch in newer version of LK functions
source("R/LatticeKrig.R")
source("R/LKrigFindLambda.R")
source("R/print.LatticeKrig.R" )
```

We begin by creating a Setup file for LatticeKrig, with only a single layer, and an a weight only  slightly above minimum (a.wght > 4). This will also create a $25 \times 25$ grid of values on our given `sDomain`.

```{r}
#uses lattice krig setup file
LKinfo<- LKrigSetup(sDomain, NC=25, nlevel=1, a.wght = 4.1,
                    NC.buffer = 0, normalize = FALSE)

#creates a couple functions, not sure what they do yet.
FUNX<- function(s){
  s[,1]
}

FUNY<- function(s){
  s[,2]
}
```




```{r}
# create matrix for fixed effects integrals
# (a linear function of coordinates in this case)
U1<- integralPolyFunction(polyGroups, M = 200)  # default is constant function 
U2<- integralPolyFunction(polyGroups,
      FUN=function(s){s[,1]}, M = 200)
U3<- integralPolyFunction(polyGroups,
      FUN=function(s){s[,2]}, M = 200)
U<- cbind( U1,U2,U3) # U can be a dense matrix
```

