---
title: "cosp_normalize.Rmd"
author: "Ryan Peterson"
date: "2024-04-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
suppressMessages(library(LatticeKrig))

suppressMessages(library( scales))
```


```{r test_data}
setwd("~/COS_LK")
load("DF_LK/DFSynthData_Base.rda")

## needs non-overlapping polyGroups and convex n-gons. Included in load data.
```

Work in Progress for updates to basisIntegral.R function

Updates include:
- Incorporate Normalization of the basis functions.
-- interpolate from the full domain down to the subset. 
- Choosing a subset of the domain around a "central" basis function.
- Replace the zGrid with the "central" basis function evaluated at integration pts. 


Fixes needed elsewhere:
- DF_Synthetic.Rmd: Change the domain to (-2,2) x (-2,2)
-- create a new surface to reflect the new domain.
-- create new file for this
-- Create convex n-gons from varonoi tesselations (see USGS code)


# Test Section

(from other code)

```{r}

w<- list( x= (1:10)/5, y=(1:15)/4)

w$z<- outer( 1:10, 1:15, FUN="+")

Center<- c( 1.5,2.5)
Scale<- .4

wNew<- list( x= (w$x - Center[1])/Scale,
             y= (w$y - Center[2])/Scale,
             z= w$z)

basisGrid<- list( x= seq(-1,1,,200), y=seq(-1,1,,200) )

wInterp<- interp.surface.grid(wNew, basisGrid)

imagePlot( w)

imagePlot( wNew)
image( wInterp, col=tim.colors(), zlim=range( c( w$z), na.rm=TRUE), add=TRUE)

imagePlot( wInterp)

```


```{r set up}
#set up Lkinfo: (testing for normalize = TRUE)
M = 200
test_Z <- outer(1:200, 1:200,  FUN="-")

sDomain<- rbind( c( -2,-2), c( 2, 2))
gridDomain <- list(x = seq(-2, 2, length.out = M),
                 y = seq(-2, 2, length.out = M))
domain_grid <- make.surface.grid(gridDomain)

#TODO: add in data? Maybe

plot(domain_grid, pch = 20, cex = 0.25)
rect(0, 0, 0.5, 0.5, border = "magenta", lwd = 3)

plot(domain_grid, pch = 20, xlim = c(0,0.5), ylim = c(0, 0.5))

#look up alpha (setting nu as 1 instead)
#a.wght = 4.01 is approx. TPS
#normalize = FALSE while we implement normalization
LKinfo <- LKrigSetup(sDomain, NC=15, nlevel=1, a.wght = 4.01,
                     NC.buffer = 0, normalize = FALSE, nu = 1)

nLevel <- LKinfo$nlevel


#TODO: plot the new data (test_Z)
gridDomain$z <- test_Z

imagePlot(gridDomain)
```

Below section

```{r bais_function_setup}

#fix this section to reflect our new choice of a basis function that exists from Lkrig.basis

reducedDomain <- rbind( c( -1,-1), c( 1,1))
gridList <- list(x = seq(-1, 1, length.out = M),
                 y = seq(-1, 1, length.out = M))

xyGrid <- make.surface.grid(gridList)

boundingBox<- rbind( c( -1,-1),
                     c( 1,-1),
                     c( 1,1),
                     c( -1, 1),
                     c( -1,-1))

dx<- gridList$x[2]- gridList$x[1]
dy<- gridList$y[2]- gridList$y[1]

zGrid <- Wendland(sqrt(xyGrid[, 1] ^ 2 + xyGrid[, 2] ^ 2),
                  dimension = 2,
                  k = 2)

stdIntegral <- sum( zGrid)*dx*dy

#remove zeros from zGrid, 

which_z <- which(zGrid != 0)

temp_xyGrid <- xyGrid[which_z,]
temp_zGrid <- zGrid[which_z]

bubblePlot(xyGrid, zGrid, highlight = FALSE)
bubblePlot(temp_xyGrid, temp_zGrid, highlight = FALSE, col = tim.colors) #nonzero
```

Normalization section

```{r normalization}
#rewrite a combination of the next 2 blocks

#directly from LKrig.basis and LKrigMarginalVariance.


l = 1 #add for l in 1:nlevel loop
basis.delta <- LKrigLatticeScales(LKinfo) 
centers <- LKrigLatticeCenters( LKinfo, Level=l ) 
PHItemp <- Radial.basis(  domain_grid, centers, basis.delta[l],
                          max.points = LKinfo$basisInfo$max.points,
                          mean.neighbor = LKinfo$basisInfo$mean.neighbor, 
                          BasisFunction = get(LKinfo$basisInfo$BasisFunction),
                          distance.type = LKinfo$distance.type,
                          verbose = FALSE)




plot(make.surface.grid(centers), pch =20)

test_wght <- LKrigNormalizeBasis(LKinfo,  Level=l, PHItemp)   #working at the moment

test_normalize <- 1/sqrt(test_wght)

test_wght_fast <- LKrigNormalizeBasisFast(LKinfo,  Level=1,  x=xyGrid)  

test_normalize_fast <- 1/sqrt(test_wght_fast)

#convert normalize to wght_matrix
wght_matrix <- matrix(unlist(test_normalize), nrow = 200, byrow = TRUE)


#plot the weight matrix 
bubblePlot(domain_grid, wght_matrix, highlight = FALSE, col = tim.colors, main = "Normalized Field") #check on title if we use this
rect(1.5, 1.5, 2, 2, border = "magenta", lwd = 3)
bubblePlot(domain_grid, wght_matrix, highlight = FALSE, col = tim.colors, xlim = c(1.5,2), ylim = c(1.5, 2))
```

# Basis Integral (Polygons)

Create some visualization for how the shifts used in `basisIntegral` functions. 
- Use the regular grid data for this. 
- Then see if we can adapt the normalization procedure to this. 

```{r polyGroups_setup}
#and viz
M <- 100
sDomain <- rbind( c( -1,-1),
                 c( 1,2))

gridDomain <- list(x = seq(-1, 1, length.out = M),
                 y = seq(-1, 2, length.out = M))
domain_grid <- make.surface.grid(gridDomain)


#repeat setup from example but with the domain for polyGroups
LKinfo2 <- LKrigSetup(sDomain, NC=15, nlevel=1, a.wght = 4.1,
                     NC.buffer = 0, normalize = FALSE, nu = 1)

N<- length(polyGroups)
cScale<- alpha(turbo(256),.5)
cTab<- color.scale(yObs_base, cScale)
layout<-setupLegend()
plot( sDomain, type="n")
for ( k in 1:N){
  polyTmp<- polyGroups[[k]]
  polygon(polyTmp, col=cTab[k], border="black", lwd=.2)
}
```


```{r polyIntegral_normal}

```



```{r}
#TODO: recreate some of the basis Integral function
#basisIntegralForEach( newpolyGroupsBig, LKinfo, M = 400, cores = 8)

#params:
## polyGroups and Lkinfo2
#using smaller M value

#from basisIntegral:

M <- 100 #make smaller if need be (default is 200) 

gridList2 <- list(x = seq(-1, 1, length.out = M),
                   y = seq(-1, 1, length.out = M))

boundingBox<- rbind( c( -1,-1),
                       c( 1,-1),
                       c( 1,1),
                       c( -1, 1),
                       c( -1,-1))

new_grid <- make.surface.grid(gridList2) #same as xyGrid below

#plot the new grid, compared to the sDomain
plot(sDomain, type = "n")
points(new_grid, pch = 20)
rect(-1, -1, 1, 1, border = "magenta", lwd = 1.5)



# setup standard integral
dx<- gridList2$x[2]- gridList2$x[1]
dy<- gridList2$y[2]- gridList2$y[1]
# integral for standard basis function.
xyGrid <- make.surface.grid(gridList2)
# values for the standard basis function on the grid
zGrid <- Wendland(sqrt(xyGrid[, 1] ^ 2 + xyGrid[, 2] ^ 2),
                    dimension = 2,
                    k = 2)
stdIntegral<- sum( zGrid)*dx*dy


nLevel <- LKinfo2$nlevel #1 for this LKInfo
N1 <- length(polyGroups)
# total number of basis functions
N2 <- LKinfo2$latticeInfo$m
# basis function scales
delta <- LKinfo2$latticeInfo$delta

#is this part needed??
integral <- NULL
J <- NULL
K <- NULL


#for (j in 1:N1) 
j <- 1
polyTmp <- (polyGroups)[[j]]


#for (L in 1:nLevel)
# get info for the L^th level of the multi-resolution basis. 
L <- 1
basisIndex <- (1:LKinfo2$latticeInfo$mLevel[L])
basisOffset<- LKinfo2$latticeInfo$offset[L]
basisScale <- LKinfo2$latticeInfo$delta[L]*LKinfo2$basisInfo$overlap
basisGridList <- LKinfo2$latticeInfo$grid[[L]]
basisCenters <- make.surface.grid(basisGridList)

#for (k in basisIndex)
inc_basis <- c(1:3, 11:13, 21:23)
#k <- 2
plot(sDomain, type = "n", xlim = c(-3.65, 1), ylim = c(-5.65, 2))
polygon(polyTmp, col="orange", border="black", lwd=.2)
rect(-1, -1, 1, 1, border = "magenta", lwd = 1.5)
rect(-1,-1,1,2, border = "darkblue", lwd = 1.5)
#standardize polygons
temp_std <- NULL
for (i in inc_basis) {
  k <- basisIndex[i]
  polyStd <- cbind((polyTmp[, 1] - basisCenters[k, 1]) / basisScale,
              (polyTmp[, 2] - basisCenters[k, 2]) / basisScale)
  temp_std <- c(temp_std, list(polyStd))
  polygon(polyStd, col=cTab[j], border="black", lwd=.2)
}


k <- 1
x_test <- c(basisCenters[k, 1], 0, basisCenters[2, 1], basisCenters[3, 1])
y_test <- c(basisCenters[k, 2], 0, basisCenters[2, 2], basisCenters[3, 2])

polyStd <- cbind((polyTmp[, 1] - basisCenters[k, 1]) / basisScale,
              (polyTmp[, 2] - basisCenters[k, 2]) / basisScale)

polyStd2 <- cbind((polyTmp[, 1] - basisCenters[2, 1]) / basisScale,
              (polyTmp[, 2] - basisCenters[2, 2]) / basisScale)

polyStd3 <- cbind((polyTmp[, 1] - basisCenters[3, 1]) / basisScale,
              (polyTmp[, 2] - basisCenters[3, 2]) / basisScale)

plot(sDomain, type = "n")
points(x_test, y_test, pch = 20)
rect(-1, -1, 1, 1, border = "magenta", lwd = 1.5)
polygon(polyTmp, col="orange", border="black", lwd=.2)
polygon(polyStd, col=cTab[j], border="black", lwd=.2)
polygon(polyStd2, col=cTab[j+2], border="black", lwd=.2)
polygon(polyStd3, col=cTab[j+4], border="black", lwd=.2)



allOutside<- in.poly(  polyStd, boundingBox)

inside <- in.poly.grid(gridList2, polyStd2)

sum(inside)
sum(zGrid[inside])  * dx * dy * basisScale ^ 2

zGrid[inside]
xyGrid[inside, 1]
xyGrid[inside, 2]

plot(xyGrid[inside, 1], xyGrid[inside, 2], pch = 20, xlim = c(0, 0.12), ylim  = c(0, 0.18))
polygon(polyStd, col=cTab[j], border="black", lwd=.2)

bubblePlot(xyGrid[inside, 1], xyGrid[inside, 2], zGrid[inside], col = tim.colors, zlim = c(0,1))


#TODO check on estimation comparison between integration points


```


```{r normaliation_data}
l = 1 #add for l in 1:nlevel loop
basis.delta <- LKrigLatticeScales(LKinfo2) 
centers <- LKrigLatticeCenters( LKinfo2, Level=l ) 
PHItemp <- Radial.basis(  domain_grid, centers, basis.delta[l],
                          max.points = LKinfo2$basisInfo$max.points,
                          mean.neighbor = LKinfo2$basisInfo$mean.neighbor, 
                          BasisFunction = get(LKinfo2$basisInfo$BasisFunction),
                          distance.type = LKinfo2$distance.type,
                          verbose = FALSE)

plot(make.surface.grid(centers), pch =20)

test_wght <- LKrigNormalizeBasis(LKinfo2,  Level=l, PHItemp)   #working at the moment

test_normalize <-  1/sqrt(test_wght)


#convert normalize to wght_matrix
wght_matrix <- matrix(unlist(test_normalize), nrow = M, byrow = TRUE)


#plot the weight matrix 
bubblePlot(domain_grid, wght_matrix, highlight = FALSE, col = tim.colors, main = "Normalization Field")

#imagePlot(gridDomain[,1], gridDomain[,2], test_normalize)

```

```{r lkrig.basis}


```


```{r test_normalization}
#TODO: apply normalization to basis integration 

wght_list <- list(x = gridDomain$x, y= gridDomain$y, z= wght_matrix)
imagePlot(wght_list, main = "Normalization Weight Field")

L <- 1

basisIndex <- (1:LKinfo2$latticeInfo$mLevel[L])
basisOffset <- LKinfo2$latticeInfo$offset[L]
basisScale <- LKinfo2$latticeInfo$delta[L]*LKinfo2$basisInfo$overlap
basisGridList <- LKinfo2$latticeInfo$grid[[L]]
basisCenters <- make.surface.grid(basisGridList)

k <- 2
#for (k in basisIndex)
x_std <- cbind((wght_list$x - basisCenters[k, 1]) / basisScale)
y_std <- cbind((wght_list$y - basisCenters[k, 2]) / basisScale)

wght_new <- list(x = x_std[,1],
                 y = y_std[,1],
                 z = wght_list$z)
image.plot(wght_new)

#interp onto the integration point grid
wght_interp <- interp.surface.grid(wght_new, gridList2)

temp_x <- wght_interp$x
temp_y <- wght_interp$y
temp_z <- wght_interp$z

image.plot(wght_interp, xlim = c(-1,1), ylim = c(-1,1))
polygon(polyStd2, col=cTab[j+2], border="black", lwd=.2) #for  k =2
#polygon(polyStd, col=cTab[j], border="black", lwd=.2) #for k =1

bubblePlot(wght_interp[inside, 1], wght_interp[inside, 2], wght_interp[inside, 3], 
           col = tim.colors, zlim = c(0.75,2))

```


```{r interp_normalize}
length(domain_grid[,1])

wght_list <- list(x = gridDomain$x, y= gridDomain$y, z= wght_matrix)



imagePlot(wght_list)
L <- 1

basisIndex <- (1:LKinfo$latticeInfo$mLevel[L])
basisOffset <- LKinfo$latticeInfo$offset[L]
basisScale <- LKinfo$latticeInfo$delta[L]*LKinfo$basisInfo$overlap
basisGridList <- LKinfo$latticeInfo$grid[[L]]
basisCenters <- make.surface.grid(basisGridList)

#check on 
k <- 85
#for (k in basisIndex)
x_std <- cbind((wght_list$x - basisCenters[k, 1]) / basisScale)
y_std <- cbind((wght_list$y - basisCenters[k, 2]) / basisScale)

wght_new <- list(x = x_std[,1],
                 y = y_std[,1],
                 z = wght_list$z)

#why are we getting NAs?
image.plot(wght_new)

wght_interp <- interp.surface.grid(wght_new, gridList)
wght_interp$z

image.plot(wght_interp, xlim = c(-1,1), ylim = c(-1,1))

range(wght_new$y)
range(wght_new$x)
image.plot(wght_new)
image( wght_interp, col=viridis(256), zlim=range( c( wght_matrix), na.rm=TRUE), add=TRUE)

image.plot(wght_interp)


```

```{r testing_NA}
length(wght_list$x)
length(basisCenters[,1])
temp_x <- wght_list$x

for (k in basisIndex) {
  x_std <- cbind((wght_list$x - basisCenters[k, 1]) / basisScale)
  y_std <- cbind((wght_list$y - basisCenters[k, 2]) / basisScale)
}


#this aligns with basisCenters
test_centers <- make.surface.grid(centers)
test.for.zero(test_centers, basisCenters)
#x_std[,1]
#(wght_list$x - basisCenters[1])/basisScale
```



```{r basis_viz}
image.plot(wght_matrix)
```

