---
title: "cosp_normalize.Rmd"
author: "Ryan Peterson"
date: "2024-04-09"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libraries}
suppressMessages(library(LatticeKrig))
```


```{r test_data}
setwd("~/COS_LK")
load("DF_LK/DFSynthData_Base.rda")

#TODO: create new synthetic data and import here:
## needs non-overlapping polyGroups and convex n-gons.
```

Work in Progress for updates to basisIntegral.R function

Updates include:
- Incorporate Normalization of the basis functions.
-- interpolate from the full domain down to the subset. 
- Choosing a subset of the domain around a "central" basis function.
- Replace the zGrid with the "central" basis function evaluated at integration pts. 


Fixes needed elsewhere:
- DF_Synthetic.Rmd: Change the domain to (-2,2) x (-2,2)
-- create a new surface to reflect the new domain.
-- create new file for this
-- Create convex n-gons from varonoi tesselations (see USGS code)


# Test Section

(from other code)

```{r}

w<- list( x= (1:10)/5, y=(1:15)/4)

w$z<- outer( 1:10, 1:15, FUN="+")

Center<- c( 1.5,2.5)
Scale<- .4

wNew<- list( x= (w$x - Center[1])/Scale,
             y= (w$y - Center[2])/Scale,
             z= w$z)

basisGrid<- list( x= seq(-1,1,,200), y=seq(-1,1,,200) )

wInterp<- interp.surface.grid(wNew, basisGrid)

imagePlot( w)

imagePlot( wNew)
image( wInterp, col=tim.colors(), zlim=range( c( w$z), na.rm=TRUE), add=TRUE)

imagePlot( wInterp)

```


```{r set up}
#set up Lkinfo: (testing for normalize = TRUE)
M = 200

sDomain<- rbind( c( -2,-2), c( 2, 2))
gridDomain <- list(x = seq(-2, 2, length.out = M),
                 y = seq(-2, 2, length.out = M))
domain_grid <- make.surface.grid(gridDomain)

#TODO: add in data? Maybe

plot(domain_grid, pch = 20, cex = 0.25)
rect(0, 0, 0.5, 0.5, border = "magenta", lwd = 3)

plot(domain_grid, pch = 20, xlim = c(0,0.5), ylim = c(0, 0.5))

#look up alpha (setting nu as 1 instead)
#a.wght = 4.01 is approx. TPS
#normalize = FALSE while we implement normalization
LKinfo <- LKrigSetup(sDomain, NC=15, nlevel=1, a.wght = 4.01,
                     NC.buffer = 0, normalize = FALSE, nu = 1)

nLevel <- LKinfo$nlevel
```

Below section

```{r bais_function_setup}

#fix this section to reflect our new choice of a basis function that exists from Lkrig.basis

reducedDomain <- rbind( c( -1,-1), c( 1,1))
gridList <- list(x = seq(-1, 1, length.out = M),
                 y = seq(-1, 1, length.out = M))

xyGrid <- make.surface.grid(gridList)

boundingBox<- rbind( c( -1,-1),
                     c( 1,-1),
                     c( 1,1),
                     c( -1, 1),
                     c( -1,-1))

dx<- gridList$x[2]- gridList$x[1]
dy<- gridList$y[2]- gridList$y[1]

zGrid <- Wendland(sqrt(xyGrid[, 1] ^ 2 + xyGrid[, 2] ^ 2),
                  dimension = 2,
                  k = 2)

stdIntegral <- sum( zGrid)*dx*dy

#remove zeros from zGrid, 

which_z <- which(zGrid != 0)

temp_xyGrid <- xyGrid[which_z,]
temp_zGrid <- zGrid[which_z]

imagePlot(xyGrid, zGrid)
bubblePlot(xyGrid, zGrid, highlight = FALSE)
bubblePlot(temp_xyGrid, temp_zGrid, highlight = FALSE, col = tim.colors) #nonzero
```

Normalization section

```{r normalization}
#rewrite a combination of the next 2 blocks

#directly from LKrig.basis and LKrigMarginalVariance.R

l = 1 #add for l in 1:nlevel loop
basis.delta <- LKrigLatticeScales(LKinfo) 
centers <- LKrigLatticeCenters( LKinfo, Level=l ) 
PHItemp <- Radial.basis(  domain_grid, centers, basis.delta[l],
                          max.points = LKinfo$basisInfo$max.points,
                          mean.neighbor = LKinfo$basisInfo$mean.neighbor, 
                          BasisFunction = get(LKinfo$basisInfo$BasisFunction),
                          distance.type = LKinfo$distance.type,
                          verbose = FALSE)




plot(make.surface.grid(centers), pch =20)

test_wght <- LKrigNormalizeBasis(LKinfo,  Level=l, PHItemp)   #working at the moment

test_normalize <- 1/sqrt(test_wght)

test_wght_fast <- LKrigNormalizeBasisFast(LKinfo,  Level=1,  x=xyGrid)  

test_normalize_fast <- 1/sqrt(test_wght_fast)

#convert normalize to wght_matrix
wght_matrix <- matrix(unlist(test_normalize), nrow = 200, byrow = TRUE)


#plot the weight matrix 
bubblePlot(domain_grid, wght_matrix, highlight = FALSE, col = tim.colors)
rect(1.5, 1.5, 2, 2, border = "magenta", lwd = 3)
bubblePlot(domain_grid, wght_matrix, highlight = FALSE, col = tim.colors, xlim = c(1.5,2), ylim = c(1.5, 2))
```



```{r interp_normalize}
length(domain_grid[,1])

wght_list <- list(x = gridDomain$x, y= gridDomain$y, z= wght_matrix)



imagePlot(wght_list)
L <- 1

basisIndex <- (1:LKinfo$latticeInfo$mLevel[L])
basisOffset <- LKinfo$latticeInfo$offset[L]
basisScale <- LKinfo$latticeInfo$delta[L]*LKinfo$basisInfo$overlap
basisGridList <- LKinfo$latticeInfo$grid[[L]]
basisCenters <- make.surface.grid(basisGridList)

for (k in basisIndex) {
  x_std <- cbind((wght_list$x - basisCenters[k, 1]) / basisScale)
  y_std <- cbind((wght_list$y - basisCenters[k, 2]) / basisScale)
}

wght_new <- list(x = x_std[,1],
                 y = y_std[,1],
                 z = wght_list$z)

#why are we getting NAs?
wght_interp <- interp.surface.grid(wght_new, gridList)
wght_interp$z

image.plot(wght_interp)

range(wght_new$y)
range(wght_new$x)
image.plot(wght_new)
image( wght_interp, col=viridis(256), zlim=range( c( wght_matrix), na.rm=TRUE), add=TRUE)
```

```{r testing_NA}
length(wght_list$x)
length(basisCenters[,1])
temp_x <- wght_list$x

for (k in basisIndex) {
  x_std <- cbind((wght_list$x - basisCenters[k, 1]) / basisScale)
  y_std <- cbind((wght_list$y - basisCenters[k, 2]) / basisScale)
}


#this aligns with basisCenters
test_centers <- make.surface.grid(centers)
test.for.zero(test_centers, basisCenters)
#x_std[,1]
#(wght_list$x - basisCenters[1])/basisScale
```



```{r basis_viz}
image.plot(wght_matrix)
```

