---
title: "cubicspline_latticekrig"
author: "Ryan Peterson"
date: "2024-09-19"
output: pdf_document
---


```{r library} 
suppressMessages(library(LatticeKrig))
```


```{r functions}
setwd("~/COS_LK")

source("DF_LK/R/bspline_poly.R")

#LK functions
source("LK_dev/ModelInterval.R")
source("LK_dev/WendlandFunction.R")
```

Goal of this document is to get `cubicSplineBasis` function working in LatticeKrig

Steps:
- Make adjustments for different "knot" locations and scale (distance between "knots") (review spatialstats project).

Additional work needed:
- Make adjustments/new `distance` function from latticekrig, to work with basis functions other than radial.


1. Begin with 1-d version
    i. Begin with 1-d version outside of LK first. 
  a. Set geometry in LatticeKrig for 1-d (in LKinfo)
    i. Through ModelInterval.R and `LKrigSetupLattice`
  b. Work through LKinfo and BasisFunction: 
    -`BasisFunction = get(LKinfo$basisInfo$BasisFunction)`
    i. Include checking behavior through LKrig.basis

2. Adapt to 2-d and tensor product
  a. Work through `Tensor.basis.R` 
  b. Plot/visualize
  
3. Get working in dev branch of LatticeKrig
  a. Run as many functions as needed here until changes are made to my branch
  b. When prepping for "final" release we will need a lot of checks.


Notes: (remove later)
 - \item{LKrigSetupLattice.LKInterval}{Creates a 1-d lattice.} (for 1-d example)
 
 
 
# 1 Test Cases without LatticeKrig

Adjustments to above outline, so that 

## 1-D Case

```{r base_nonLK}
#setup up grid/knots at normal intervals (figure our delta later)
x_grid <- seq(0, 6, 0.01)
nc <- 8 #number of interior grid points
buffer <- 1
overlap <- 1  #fixed

rangeLocations <-  range(x_grid)

grid.info <- list( xmin = rangeLocations[1], xmax= rangeLocations[2], 
                   range = rbind(rangeLocations))
delta <- ( grid.info$xmax - grid.info$xmin ) / ( nc - 1 )

buffer.width <- buffer * delta
grid.list <- list(x = seq(grid.info$xmin - buffer.width, 
            grid.info$xmax + buffer.width, delta) ) #centers

class( grid.list) <- "gridList"  #(add if needed)

#with nlevel > 1 mLevel[j]
mLevel <- length(grid.list$x)
offset <- as.integer(c(0, cumsum(mLevel)))

basis.delta <- (delta*overlap)


xgrid_bsplines <- matrix(NA, nrow = length(x_grid))
for(j in grid.list$x){
  xgrid_bsplines <- cbind(xgrid_bsplines, cubicSplineBasis(((x_grid-j)/basis.delta)))
}
xgrid_bsplines <- xgrid_bsplines[,-1]
  

test_grid <- seq(min(grid.list$x), max(grid.list$x), 0.01)

test_bsplines <- matrix(NA, nrow = length(test_grid))
for(j in grid.list$x){
  test_bsplines <- cbind(test_bsplines, cubicSplineBasis(((test_grid-j)/basis.delta)))
}
test_bsplines <- test_bsplines[,-1]

#rowSums(test_bsplines)

plot(range(grid.list$x), c(0,1), type = "n", xlab = "x", ylab = "")
abline(v = grid.list$x, lty = 2)
abline(v = rangeLocations, lty = 2, lwd = 2)
matlines(test_grid, test_bsplines, ylim = c(0,1), lty = 1)
matlines(x_grid, xgrid_bsplines, ylim = c(0,1), lty = 1, lwd = 2.2)
lines(test_grid, rowSums(test_bsplines), col = "darkred", lwd = 2)
```

# 2-D Tensor Product

(Look up rectangular geometry setup code)

```{r}
x_grid <- seq(0, 6, 0.01)
y_grid <- seq(0, 6, 0.01)
nc <- 9  #number of grid points
buffer <- 1
overlap <- 1  #fixed

range_xgrid <-  range(x_grid)
range_ygrid <-  range(y_grid)

grid.info_x <- list( xmin = range_xgrid[1], xmax= range_xgrid[2], 
                   range = rbind(range_xgrid))
grid.info_y <- list( ymin = range_ygrid[1], ymax= range_ygrid[2], 
                   range = rbind(range_ygrid))

delta <- ( grid.info_x$xmax - grid.info_x$xmin ) / ( nc - 1 )
buffer.width <- buffer * delta

#lattice centers
grid.list_full <- list(x = seq(grid.info_x$xmin - buffer.width, 
            grid.info_x$xmax + buffer.width, delta),
            y = seq(grid.info_y$ymin - buffer.width, 
            grid.info_y$ymax + buffer.width, delta)) #centers

class( grid.list_full) <- "gridList"  

basis.delta <- (delta*overlap)

#b-spline test
test.grid_x <- seq(min(grid.list_full$x-1), max(grid.list_full$x+1), 0.01)
test.grid_y <- seq(min(grid.list_full$y-1), max(grid.list_full$y+1), 0.01)

test_bsplines_x <- matrix(NA, nrow = length(test.grid_x))
for(j in grid.list_full$x){
  test_bsplines_x <- cbind(test_bsplines_x, cubicSplineBasis(((test.grid_x-j)/basis.delta)))
}
test_bsplines_y <- matrix(NA, nrow = length(test.grid_y))
for(k in grid.list_full$y){
  test_bsplines_y <- cbind(test_bsplines_y, cubicSplineBasis(((test.grid_y-k)/basis.delta)))
}

test_bsplines_x <- test_bsplines_x[ ,-1]
test_bsplines_y <- test_bsplines_y[ ,-1]

tensor_list <- list()
for (i in 1:length(grid.list_full$x)) {
  for (k in 1:length(grid.list_full$y)) {
    phi_x <- test_bsplines_x[ ,i]
    phi_y <- test_bsplines_y[ ,k]
    tensor_xy <- outer(phi_x, phi_y, "*")
    tensor_list[[paste0("x",i,"y",k)]] <- tensor_xy
  }
}

tensor_sum <-  Reduce("+", tensor_list)
xy_surface <- make.surface.grid( list(x=test.grid_x, y= test.grid_y) )
image.plot( as.surface(xy_surface, tensor_sum))
rect(0,0,6,6, col = NA, border = "magenta", lwd = 2)

```

```{r}
length(which(tensor_sum == 1))
xy_surface[187,187]
```


