---
title: "Footpring_dataCleaning"
author: "Ryan Peterson"
date: "2024-03-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r library}
suppressMessages(library(lubridate))
suppressMessages(library(fields))
suppressMessages(library(ncdf4)) #for .nc file type
suppressMessages(library(raster))
suppressMessages(library(viridis))
suppressMessages(library( scales))

```


Bounding region, use this to restrict to only the Australia and Maritime SE Asia

```{r bounding_Region}
Lat_min <- -50
Lat_max <- 10
Lon_min <- 90
Lon_max <- 180
bounds <- data.frame(Lon_min, Lon_max, Lat_min, Lat_max )

full_domain <- rbind(c(90,-50), 
                     c(180,10))
```


# TROPOMI Data

TODO: add in multiple data sets to get full coverage of the region.

```{r data_import}
#-----testing-----##

#test import

##first

file.exists('Data/S5P_OFFL_L2__CO_____20221226T015719_20221226T033848_26952_03_020400_20221229T072250.nc')


this.nc <- nc_open('Data/S5P_OFFL_L2__CO_____20221226T015719_20221226T033848_26952_03_020400_20221229T072250.nc')

##

file.exists('Data/S5P_OFFL_L2__CO_____20221226T033848_20221226T052018_26953_03_020400_20221229T072319.nc')

this2.nc <- nc_open('Data/S5P_OFFL_L2__CO_____20221226T033848_20221226T052018_26953_03_020400_20221229T072319.nc')

##

file.exists('Data/S5P_OFFL_L2__CO_____20221226T052018_20221226T070148_26954_03_020400_20221229T072348.nc')

this3.nc <- nc_open('Data/S5P_OFFL_L2__CO_____20221226T052018_20221226T070148_26954_03_020400_20221229T072348.nc')


#last

file.exists('Data/S5P_OFFL_L2__CO_____20221226T070148_20221226T084319_26955_03_020400_20221229T072418.nc')

this4.nc <- nc_open('Data/S5P_OFFL_L2__CO_____20221226T070148_20221226T084319_26955_03_020400_20221229T072418.nc')

# test_val.grid <- this.nc[["dim"]][[]]
```
Total Column CO and lon/lat bounds

```{r}
#data 1 (pass 1)
total_col_co <- ncvar_get(this.nc, "PRODUCT/carbonmonoxide_total_column") #lots of NA
test_lonbond <- ncvar_get(this.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/longitude_bounds")
test_latbond <- ncvar_get(this.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/latitude_bounds")

#data 2 (pass 2)
total_col_co_2 <- ncvar_get(this2.nc, "PRODUCT/carbonmonoxide_total_column") #lots of NA
test_lonbond_2 <- ncvar_get(this2.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/longitude_bounds")
test_latbond_2 <- ncvar_get(this2.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/latitude_bounds")

#data 3 (pass 3)
total_col_co_3 <- ncvar_get(this3.nc, "PRODUCT/carbonmonoxide_total_column") #lots of NA
test_lonbond_3 <- ncvar_get(this3.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/longitude_bounds")
test_latbond_3 <- ncvar_get(this3.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/latitude_bounds")

#data 4 (pass 4)
total_col_co_4 <- ncvar_get(this4.nc, "PRODUCT/carbonmonoxide_total_column") #lots of NA
total_col_co_4 <- ncvar_get(this4.nc, "PRODUCT/carbonmonoxide_total_column_corrected") #using delined data
test_lonbond_4 <- ncvar_get(this4.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/longitude_bounds")
test_latbond_4 <- ncvar_get(this4.nc, "PRODUCT/SUPPORT_DATA/GEOLOCATIONS/latitude_bounds")
```



for below matrices they are 215 x 4172, where
ground pixel 215: The dimension perpendicular to the flight direction.
scanline 4172: The The dimension that indicates the flight direction.

latitude in CO____/PRODUCT
Description: The latitude of the pixel centers of the ground pixels in the data. Latitude, longitude
coordinates for the ground pixel center and the ground pixel corners are calculated at the
WGS84 ellipsoid.

longitude in CO____/PRODUCT
Description: The longitude of the pixel centers of the ground pixels in the data. Latitude, longitude
coordinates for the ground pixel center and the ground pixel corners are calculated at the
WGS84 ellipsoid.

carbonmonoxide_total_column in CO____/PRODUCT
Description: Vertically integrated CO column density


From this and the bounds we get a centroid for each pixel as latitude,longitude.
-Then the bounds are defined from latitude_bounds in CO____/PRODUCT/SUPPORT_DATA/GEOLOCATIONS and longitude_bounds in CO____/PRODUCT/SUPPORT_DATA/GEOLOCATIONS. 

```{r function_trop_polygons}
tropomi_polygons <- function(total_col_co, test_lonbond, test_latbond, bounds){
  
  #lots of manual movements (fix later)
  trop_lonbond_loc1 <- test_lonbond[1,,]
  trop_lonbond_loc2 <- test_lonbond[2,,]
  trop_lonbond_loc3 <- test_lonbond[3,,]
  trop_lonbond_loc4 <- test_lonbond[4,,]
  
  trop_latbond_loc1 <- test_latbond[1,,]
  trop_latbond_loc2 <- test_latbond[2,,]
  trop_latbond_loc3 <- test_latbond[3,,]
  trop_latbond_loc4 <- test_latbond[4,,]
  
  #get locations in matrix of Non_NA
  trop_col_co <- total_col_co[which(!is.na(total_col_co))]
  
  #select for pixel locations not associated with NA
  trop_lonbond_loc1 <- trop_lonbond_loc1[which(!is.na(total_col_co))]
  trop_lonbond_loc2 <- trop_lonbond_loc2[which(!is.na(total_col_co))]
  trop_lonbond_loc3 <- trop_lonbond_loc3[which(!is.na(total_col_co))]
  trop_lonbond_loc4 <- trop_lonbond_loc4[which(!is.na(total_col_co))]
  
  trop_latbond_loc1 <- trop_latbond_loc1[which(!is.na(total_col_co))]
  trop_latbond_loc2 <- trop_latbond_loc2[which(!is.na(total_col_co))]
  trop_latbond_loc3 <- trop_latbond_loc3[which(!is.na(total_col_co))]
  trop_latbond_loc4 <- trop_latbond_loc4[which(!is.na(total_col_co))]
  
  #delete blocks outsides of bounds
  N <- length(trop_col_co)
  delete_list <- NA
  for (i in 1:N) {
    if(trop_lonbond_loc2[i] < bounds$Lon_min | trop_latbond_loc1[i] < bounds$Lat_min | 
       trop_latbond_loc1[i] >  bounds$Lat_max){
      delete_list<- c( delete_list,i)
    }
  }
  delete_list <- delete_list[-1]
  
  #delete step
  trop_col_co <- trop_col_co[-delete_list]

  trop_lonbond_loc1 <- trop_lonbond_loc1[-delete_list]
  trop_lonbond_loc2 <- trop_lonbond_loc2[-delete_list]
  trop_lonbond_loc3 <- trop_lonbond_loc3[-delete_list]
  trop_lonbond_loc4 <- trop_lonbond_loc4[-delete_list]
  
  trop_latbond_loc1 <- trop_latbond_loc1[-delete_list]
  trop_latbond_loc2 <- trop_latbond_loc2[-delete_list]
  trop_latbond_loc3 <- trop_latbond_loc3[-delete_list]
  trop_latbond_loc4 <- trop_latbond_loc4[-delete_list]
  
  #get polygroups
  polyGroup <- NA
  N <- length(trop_lonbond_loc1)
  #N <- 50000
  for(i in 1:N){
    newPoly <- rbind(c(trop_lonbond_loc1[i], trop_latbond_loc1[i]),
                     c(trop_lonbond_loc2[i], trop_latbond_loc2[i]),
                     c(trop_lonbond_loc3[i], trop_latbond_loc3[i]),
                     c(trop_lonbond_loc4[i], trop_latbond_loc4[i]),
                     c(trop_lonbond_loc1[i], trop_latbond_loc1[i]))
    polyGroup<- c( polyGroup,list(newPoly))
  }
  
  polyGroup <- polyGroup[-1]
  test_col <- trop_col_co[1:N]
  return(list(polygons = polyGroup, data = test_col))
}
```


```{r test_function}
out <- tropomi_polygons(total_col_co, test_lonbond, test_latbond, bounds)
```


```{r true_cleaning}


trop_lonbond_loc1 <- test_lonbond[1,,]
trop_lonbond_loc2 <- test_lonbond[2,,]
trop_lonbond_loc3 <- test_lonbond[3,,]
trop_lonbond_loc4 <- test_lonbond[4,,]

trop_latbond_loc1 <- test_latbond[1,,]
trop_latbond_loc2 <- test_latbond[2,,]
trop_latbond_loc3 <- test_latbond[3,,]
trop_latbond_loc4 <- test_latbond[4,,]

#get locations in matrix of non-NA
trop_col_co <- total_col_co[which(!is.na(total_col_co))]

trop_lonbond_loc1 <- trop_lonbond_loc1[which(!is.na(total_col_co))]
trop_lonbond_loc2 <- trop_lonbond_loc2[which(!is.na(total_col_co))]
trop_lonbond_loc3 <- trop_lonbond_loc3[which(!is.na(total_col_co))]
trop_lonbond_loc4 <- trop_lonbond_loc4[which(!is.na(total_col_co))]

trop_latbond_loc1 <- trop_latbond_loc1[which(!is.na(total_col_co))]
trop_latbond_loc2 <- trop_latbond_loc2[which(!is.na(total_col_co))]
trop_latbond_loc3 <- trop_latbond_loc3[which(!is.na(total_col_co))]
trop_latbond_loc4 <- trop_latbond_loc4[which(!is.na(total_col_co))]
```


```{r check_bounds}
N <- length(trop_col_co)
delete_list <- NA
for (i in 1:N) {
  if(trop_lonbond_loc2[i] < Lon_min | trop_latbond_loc1[i] < Lat_min | 
     trop_latbond_loc1[i] >  Lat_max){
    delete_list<- c( delete_list,i)
  }
}
delete_list <- delete_list[-1]
```


```{r delete_outside values}
trop_col_co <- trop_col_co[-delete_list]

trop_lonbond_loc1 <- trop_lonbond_loc1[-delete_list]
trop_lonbond_loc2 <- trop_lonbond_loc2[-delete_list]
trop_lonbond_loc3 <- trop_lonbond_loc3[-delete_list]
trop_lonbond_loc4 <- trop_lonbond_loc4[-delete_list]

trop_latbond_loc1 <- trop_latbond_loc1[-delete_list]
trop_latbond_loc2 <- trop_latbond_loc2[-delete_list]
trop_latbond_loc3 <- trop_latbond_loc3[-delete_list]
trop_latbond_loc4 <- trop_latbond_loc4[-delete_list]
```


```{r}
#test with 50k


polyGroup <- NA
N <- length(trop_lonbond_loc1)
#N <- 50000
for(i in 1:N){
  newPoly <- rbind(c(trop_lonbond_loc1[i], trop_latbond_loc1[i]),
                   c(trop_lonbond_loc2[i], trop_latbond_loc2[i]),
                   c(trop_lonbond_loc3[i], trop_latbond_loc3[i]),
                   c(trop_lonbond_loc4[i], trop_latbond_loc4[i]),
                   c(trop_lonbond_loc1[i], trop_latbond_loc1[i]))
  polyGroup<- c( polyGroup,list(newPoly))
}

polyGroup <- polyGroup[-1]
test_col <- trop_col_co[1:N]
```


```{r}
out <- tropomi_polygons(total_col_co, test_lonbond, test_latbond, bounds)
out2 <- tropomi_polygons(total_col_co_2, test_lonbond_2, test_latbond_2, bounds)


```


```{r}
out3 <- tropomi_polygons(total_col_co_3, test_lonbond_3, test_latbond_3, bounds)
out4 <- tropomi_polygons(total_col_co_4, test_lonbond_4, test_latbond_4, bounds)

polyGroups <- c(out$polygons, out2$polygons, out3$polygons, out4$polygons)
col_co <- c(out$data, out2$data, out4$data, out4$data)
covert_col_co <- col_co *6.022140857e+19
scale_lim <- range(covert_col_co)
zero_vals <- which(covert_col_co <= 0)
zero_vals

covert_col_co <- covert_col_co[-zero_vals]
polyGroups <- polyGroups[-zero_vals]
```





lets try this but with lat lon centers, if this is good enough we can run with it


# MOPITT Level 3

```{r}

file.exists('Data/MOP02J-20221226-L2V19.10.3.he5')


mopitt.nc <- nc_open('Data/MOP02J-20221226-L2V19.10.3.he5')
```

Grid:
"HDFEOS/GRIDS/MOP03/Data Fields/NumberofPixelsDay"

```{r}
grid_mopitt <-  ncvar_get(mopitt.nc, "HDFEOS/GRIDS/MOP03/Data Fields/NumberofPixelsDay")
```



# MOPITT Level 2

```{r}

file.exists('Data/MOP02J-20221226-L2V19.10.3.he5')


mopitt.nc <- nc_open('Data/MOP02J-20221226-L2V19.10.3.he5')
```

"HDFEOS/GRIDS/MOP03/Data Fields/RetrievedCOTotalColumnDay" correct this
Total Column day

```{r}
mopitt_co <- ncvar_get(mopitt.nc, "HDFEOS/SWATHS/MOP02/Data Fields/RetrievedCOTotalColumn")
mopitt_co <- mopitt_co[1,] #select for parameter
```

Pixel "centers"
pixel are 22x22km
-pixel polygon corners are 


"HDFEOS/SWATHS/MOP02/Geolocation Fields/Latitude"
Latitude
"HDFEOS/GRIDS/MOP03/Data Fields/Longitude" Correct this
Longitude

```{r}
test_mopitt_lat <-  ncvar_get(mopitt.nc, "HDFEOS/SWATHS/MOP02/Geolocation Fields/Latitude")
test_mopitt_lon <-  ncvar_get(mopitt.nc, "HDFEOS/SWATHS/MOP02/Geolocation Fields/Longitude")
```

```{r NA_clean}
mopitt_col_co <- mopitt_co[which(!is.na(mopitt_co))]

mopitt_lon <- test_mopitt_lon[which(!is.na(mopitt_co))]
mopitt_lat <- test_mopitt_lat[which(!is.na(mopitt_co))]
```

```{r boundary_check}
N <- length(mopitt_col_co)
delete_list_2 <- NA
for (j in 1:N) {
  if(mopitt_lon[j] < bounds$Lon_min | mopitt_lat[j] < bounds$Lat_min | 
     mopitt_lat[j] > bounds$Lat_max){
      delete_list_2<- c( delete_list_2,j)
  }
}
delete_list_2 <- delete_list_2[-1]
```

```{r}
mopitt_col_co <- mopitt_col_co[-delete_list_2]

mopitt_lon <- mopitt_lon[-delete_list_2]
mopitt_lat <- mopitt_lat[-delete_list_2]
```


```{r}
lat_kmdeg <- 1 / 110.574 #lat km in degrees
1 / (111.320 * cos(mopitt_lat[20] * pi / 180)) #lon km in degrees
```


```{r}
lat_kmdeg <- 1 / 110.574 #lat km in degrees
lat_adj <- 11 * lat_kmdeg

N <- length(mopitt_lon)
mop_lon_loc1 <- NA
mop_lon_loc2 <- NA
mop_lon_loc3 <- NA
mop_lon_loc4 <- NA
  
mop_lat_loc1 <- NA
mop_lat_loc2 <- NA
mop_lat_loc3 <- NA
mop_lat_loc4 <- NA

for(k in 1:N){
  temp_lon_adj <- 11 * (1 / (111.320 * cos(mopitt_lat[k] * pi / 180)))
  mop_lon_loc1 <- c(mop_lon_loc1, mopitt_lon[k]-temp_lon_adj)
  mop_lon_loc2 <- c(mop_lon_loc2, mopitt_lon[k]+temp_lon_adj)
  mop_lon_loc3 <- c(mop_lon_loc3, mopitt_lon[k]+temp_lon_adj)
  mop_lon_loc4 <- c(mop_lon_loc4, mopitt_lon[k]-temp_lon_adj)
  
  mop_lat_loc1 <- c(mop_lat_loc1, mopitt_lat[k]-lat_adj)
  mop_lat_loc2 <- c(mop_lat_loc2, mopitt_lat[k]-lat_adj)
  mop_lat_loc3 <- c(mop_lat_loc3, mopitt_lat[k]+lat_adj)
  mop_lat_loc4 <- c(mop_lat_loc4, mopitt_lat[k]+lat_adj)
}

mop_lon_loc1 <- mop_lon_loc1[-1]
mop_lon_loc2 <- mop_lon_loc2[-1]
mop_lon_loc3 <- mop_lon_loc3[-1]
mop_lon_loc4 <- mop_lon_loc4[-1]
  
mop_lat_loc1 <- mop_lat_loc1[-1]
mop_lat_loc2 <- mop_lat_loc2[-1]
mop_lat_loc3 <- mop_lat_loc3[-1]
mop_lat_loc4 <- mop_lat_loc4[-1]

```


```{r}
polyGroup_mop <- NA
N <- length(mop_lon_loc1)
#N <- 50000
for(i in 1:N){
  newPoly <- rbind(c(mop_lon_loc1[i], mop_lat_loc1[i]),
                   c(mop_lon_loc2[i], mop_lat_loc2[i]),
                   c(mop_lon_loc3[i], mop_lat_loc3[i]),
                   c(mop_lon_loc4[i], mop_lat_loc4[i]),
                   c(mop_lon_loc1[i], mop_lat_loc1[i]))
  polyGroup_mop<- c( polyGroup_mop,list(newPoly))
}


polyGroup_mop <- polyGroup_mop[-1]
test_col_mop <- mopitt_col_co[1:N]

test_col_mop <- test_col_mop[order(mopitt_col_co)]
polyGroup_mop <- polyGroup_mop[order(mopitt_col_co)]
range(test_col_mop)
```


```{r}
fixed_order <- order(covert_col_co)
covert_col_co <- covert_col_co[fixed_order]
polyGroups <- polyGroups[fixed_order]
```


```{r}
trop_range <- range(covert_col_co)
mop_range <- range(test_col_mop)
trop_range
mop_range
min(trop_range, mop_range)
max(trop_range, mop_range)

scale_lim <- c(min(trop_range, mop_range), max(trop_range, mop_range))

save( covert_col_co, test_col_mop,
      polyGroups, polyGroup_mop,
    file =  "Trop_Mop_viz.rda")
```

```{r AUS_boundaries}
#set NE boundary points
NE_maxLon <- 155
NE_minLon <- 134
NE_maxLat <- -10
NE_minLat <- -25

#set SE boundary points
SE_maxLon <- 155
SE_minLon <- 134
SE_maxLat <- -25
SE_minLat <- -48
```


```{r tropomi_viz}
N<- length(polyGroups)
cScale<- alpha(turbo(256),.75)
cTab<- color.scale(covert_col_co, cScale, zlim = scale_lim)
layout <- setupLegend()
plot( full_domain, type="n",xlab = "Longitude", ylab = "Latitude",
      main = "TROPOMI - Total Column CO")
for ( k in 1:N){
  polyTmp<- polyGroups[[k]]
  polygon(polyTmp, col=cTab[k], border=NA)
}
world(add = TRUE)
rect(NE_minLon, NE_minLat, NE_maxLon, NE_maxLat, border = "magenta", lwd = 3)
#text(143, -14, "NE Aus", col = "magenta")
rect(SE_minLon, SE_minLat, SE_maxLon, SE_maxLat, border = "magenta", lwd = 3)
#text(143, -30, "SE Aus", col = "magenta")
addLegend(layout, col=cTab, zlim = scale_lim, 
          legend.args=list( text="CO [mol/cm^2]", cex=1, side=1, line=1))
#TODO: add z range on the right
```

```{r mopitt_viz}
N<- length(polyGroup_mop)
cScale<- alpha(turbo(256),.75)
cTab <- color.scale(test_col_mop, cScale,zlim = scale_lim)
layout<-setupLegend()
plot( full_domain, type="n", xlab = "Longitude", ylab = "Latitude", 
      main = "MOPITT - Total Column CO")
for ( k in 1:N){
  polyTmp<- polyGroup_mop[[k]]
  polygon(polyTmp, col=cTab[k], border=NA)
}
world(add = TRUE)
rect(NE_minLon, NE_minLat, NE_maxLon, NE_maxLat, border = "magenta", lwd = 3)
#text(143, -14, "NE Aus", col = "magenta")
rect(SE_minLon, SE_minLat, SE_maxLon, SE_maxLat, border = "magenta", lwd = 3)
#text(143, -30, "SE Aus", col = "magenta")
addLegend(layout, col=cTab, zlim = scale_lim, 
          legend.args=list( text="CO [mol/cm^2]", cex=1, side=1, line=1))
#mtext("CO [mol/cm^2]", side = 1, adj = 1.25)
```

```{r}
set.panel(1,2)
N<- length(polyGroup_mop)
cScale<- alpha(turbo(256),.75)
cTab <- color.scale(test_col_mop, cScale,zlim = scale_lim)
layout<-setupLegend()
plot( full_domain, type="n", xlab = "Longitude", ylab = "Latitude", 
      main = "MOPITT - Total Column CO")
for ( k in 1:N){
  polyTmp<- polyGroup_mop[[k]]
  polygon(polyTmp, col=cTab[k], border=NA)
}
world(add = TRUE)
addLegend(layout, col=cTab, zlim = scale_lim, 
          legend.args=list( text="CO [mol/cm^2]", cex=1, side=1, line=1))

N<- length(polyGroups)
cScale<- alpha(turbo(256),.75)
cTab<- color.scale(covert_col_co, cScale, zlim = scale_lim)
layout<-setupLegend()
plot( full_domain, type="n",xlab = "Longitude", ylab = "Latitude",
      main = "TROPOMI - Total Column CO")
for ( k in 1:N){
  polyTmp<- polyGroups[[k]]
  polygon(polyTmp, col=cTab[k], border=NA)
}
world(add = TRUE)
addLegend(layout, col=cTab, zlim = scale_lim , 
          legend.args=list( text="CO [mol/cm^2]", cex=1, side=1, line=1))
```

