---
title: "LK Update Notes"
author: "Ryan Peterson"
date: "2025-01-16"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r library}
suppressMessages(library(LatticeKrig))
```

```{r test_data}
setwd("~/COS_LK")

#data:
load(file = "DF_LK/synthetic_data.rda") 

synth_df <- data.frame(sGrid, look_stat)
LKinfo_base <- LKinfo_stat

#sample from synthetic data
n <- 750
set.seed(350)
sub1_df <- synth_df[sample(nrow(synth_df), n, replace = FALSE), ]

#sample plot
bubblePlot(sub1_df$x, sub1_df$y, sub1_df$look_stat, 
           col = tim.colors, main = "Synth Data - Sample Points")

```

```{r imported_functions}
setwd("~/COS_LK")
#functions (custom)
source("DF_LK/R/bspline_poly.R") #cubic b-splines basis function

#functions (LK_dev)
source("LK_dev/LKDistComponents.R")
source("LK_dev/Tensor.basis.R")
source("LK_dev/LKrig.basis.R")

#source("LK_dev/ModelRectangle.R")
```


Code to test implementation of the the basis function directly into LKrig

Needed function/files (and notes):
(work backwards?, figuring out where the basis function enters into the code.)
 
- LatticeKrig(): for the mle (replace with MLE later), adds mle to LKrig()
-- Do we need to have an updated basis function here?

- LKrig(): main function for latticekrig

- LKrigSetup(): essential for setting up basically everything in LK


# LKrigSetup()

Sets LKinfo generally

TODO: flesh this out later

# LKrig()

Parameters: (x, y,...)

Line 44: LKinfo <- do.call("LKrigSetup", c(list(x = x ), list(...),...)
Line 48: LKinfo<- do.call("LKinfoUpdate", c(list(LKinfo=LKinfo), list(...)) )

TODO: check LKinfoUpdate to see if we will have an issue with new features added to LKinfo. 

Recall do.call 

Line 89-94:  object<- createLKrigObject( x, y, weights, Z, X, U, LKinfo,...
- Used to create objects used in later function calls within
- Note X in above function is for the basis functions.

Line 121: wX<- LKrigMakewX( object, verbose=verbose)
 

## LKrigMakewX

Sets up wX with known or unknown wX (weighted basis function matrix)

Line: 30-31  wX <- diag.spam(sqrt(object\$weights)) %*% LKrig.basis(object\$x, object\$LKinfo, verbose = FALSE)
- This is from the else (no object\$X) and uses LKrig.basis to set-up object\$X.

## LKrig.basis

(locally: COS_LK/LK_dev)

We use the tensor type with modifications:

Lines 86-91: PHItemp <- Tensor.basis(  x1, centers, basis.delta[l],...)
- There are a lot of LKinfo setting that are used before (and in this function call), they are:

LKinfo\$ :
- nlevel (number of levels)
- normalize (not needed)
- distance.type (used redundantly in Tensor.basis function call)
basisInfo\$ :
- BasisType=="Tensor" (sets type of basis function we are using)
- max.points ()
- mean.neighbor ()
- BasisFunction () 


We also use 
Line 63: basis.delta <- LKrigLatticeScales(LKinfo) 
- Call to this other function to get basis.delta (distance between nodes)


### Tensor.basis

Default Parameters: (x1, centers, basis.delta,
                     max.points = NULL, 
                  mean.neighbor = 50,
                  BasisFunction = "WendlandFunction", 
                  distance.type = "Euclidean"

passes basically everything to LKrigDistance, include the components = TRUE
(Radial.basis uses components = FALSE)

Line 42:  out<- LKrigDistance( x1, centers,...)
- This gets distances to each node center

Line 53:  temp <- do.call( BasisFunction, list(d=out$ra[,1]))
- evaluates the basis function

### LKrigDistance(Components)

Begin here to check changes from radial to tensor basis function.

Line 39-41: LKDistComponents(x1, x2, delta,
	     max.points = max.points, mean.neighbor = mean.neighbor,
	     distance.type = distance.type)
- Uses LKDistComponents is components = TRUE


Parameters: (  x1, x2, delta,...
- x1: data locations
- x2: lattice centers

# Testing Section

## LKinfo Changes

This is a list of things that need to be done or changed to LKinfo

overlap = 2

```{r lkinfo_changes}
#overlap for cubic b-spline is 2
LKinfo_stat$basisInfo$overlap <- 2

#change basis function type
LKinfo_stat$basisInfo$BasisFunction <- "CubicBSpline"
```


## LKrig.basis Test/Change

Testing and changing LKrig.basis() and other functions to allow for the new basis function.

Beginning at the bottom, working up

1. LKrigDistance()

With components = TRUE uses LKDistComponents()

Using code from Tensor.basis and LKrig.basis to setup necessary variables.

```{r lkdist_test}
#testing LKDistComponents()

#setup
#using LKinfo_stat as our LKinfo
#repeated again below when testing Tensor.basis and LKrig.basis
nlevel <- LKinfo_stat$nlevel #number of levels
distance.type <- LKinfo_stat$distance.type
basis.delta <- LKrigLatticeScales(LKinfo_stat)
centers <- LKrigLatticeCenters(LKinfo_stat, Level=1)

SE_coords <- cbind(sub1_df$x, sub1_df$y) #x1 (locations)
mean.neighbor <- 50


dist_test <- LKrigDistance( SE_coords, centers,
                             delta =  basis.delta,
                        max.points = max.points,
                     mean.neighbor = mean.neighbor,
                     distance.type = distance.type,
                        components = TRUE)    

#dist_test <- LKDistComponents(SE_coords, centers, basis.delta)
```


2. Tensor.basis()

```{r tensor_test}
#Tensor.basis setup
x1 <- SE_coords
max.points = NULL

dimension <- ncol(x1)
n1 <- nrow(x1)
if (is.null(max.points)) {
    Nmax <- n1 * mean.neighbor
} else {
    Nmax <- max.points
}

#adjusting dist_Test for scale
dist_test$ra <- dist_test$ra/basis.delta

#basis function call
BasisFunction <- "WendlandFunction"

wendland_out <- do.call( BasisFunction, list(d=dist_test$ra[,1]))

BasisTest <- "CubicBSpline"

bspline_out <- do.call( BasisTest, list(d=dist_test$ra[,1]))

if (dimension > 1) {
    for (j in (2:dimension)) {
        wendland_out <- wendland_out * do.call( BasisFunction, list( d=dist_test$ra[,j]))
        bspline_out <-  bspline_out * do.call( BasisTest, list(d=dist_test$ra[,j]))
    }
}



```

```{r full_tensor_test}
l <- 1
x1 <- SE_coords
PHItemp <- Tensor.basis(  x1, centers, basis.delta[l],
                                max.points = LKinfo_stat$basisInfo$max.points,
                             mean.neighbor = LKinfo_stat$basisInfo$mean.neighbor, 
                             BasisFunction = get(LKinfo_stat$basisInfo$BasisFunction),
                             distance.type = LKinfo_stat$distance.type)
```


3. LKrig.basis()

Add in some internals from LKrig.basis to update the phi (basis function) 

```{r lkrig.basis_test}

#updates to LKinfo
LKinfo_stat$basisInfo$BasisType <- "Tensor"
LKinfo_stat$normalize <- FALSE
LKinfo_stat$BasisFunction <- "CubicBSpline"
LKinfo_stat$BasisType <- "Tensor"

#checking on alphaObject, is null in this 
#Lines 139-145
if (is.null( LKinfo_stat$alphaObject[[l]]) ){
  wght <- LKinfo_stat$alpha[[l]]
} else{
# spatially varying alpha extension         
  wght <- LKinfo_stat$alpha[[l]]*c(predict(LKinfo_stat$alphaObject[[l]], x1))
}


# rho may not be provided when estimating a model 
#Lines 167-173
if (!is.null( LKinfo_stat$rho.object) ) {
  wght <- c(predict(LKinfo_stat$rho.object, x1))
} else{
  wght<- ifelse( is.na( LKinfo_stat$rho), 1.0, LKinfo_stat$rho )
}

#update PHI temp with rho
PHI <-sqrt(wght)*PHItemp
```

4. LKrig()

Test using LKrig but pass in X=PHI

```{r lkrig}
SE_coords <- cbind(sub1_df$x, sub1_df$y)
SE_data <- sub1_df$look_stat

out <- LKrig(SE_coords, SE_data, LKinfo = LKinfo_stat, X = PHI)
out2 <-  LKrig(SE_coords, SE_data, LKinfo = LKinfo_base)

```


```{r predict}

M <- 48
gridList_new <- list( x= seq( -1,1,length.out= M),
                 y= seq( -1,1,length.out= M) )
sGrid_new <- make.surface.grid(gridList_new)

lk_hat <- predictSurface(out, gridList_new)

#TODO: png output
image.plot(lk_hat)

setwd("~/COS_LK")
png(filename = "pred_new2.png", width = 1800, height = 1800, res = 300)
image.plot(lk_hat, 
           xlab = "x", ylab = "y", main = "Predicted - Cubic B-Spline")
dev.off()
```

```{r test}
#using base wendland, without MLE
lk_hat2 <- predictSurface(out2, gridList_new)

#TODO: png output
image.plot(lk_hat2)

setwd("~/COS_LK")
png(filename = "pred_base2.png", width = 1800, height = 1800, res = 300)
image.plot(lk_hat2, 
           xlab = "x", ylab = "y", main = "Predicted - Wendland")
dev.off()
```

```{r}
#TODO: get side by side comparisons and differencesof Wendland, Cubic B-spline, etc (maybe compare )
```


```{r}
#TODO: get MLE working for this
## Integrate the new basis function into a local version of LatticeKrig
## Redo the "base" surface with new function (e.g. tensorprod_surface.png) 


#everything below needs work.
out2 <- LatticeKrig(SE_coords, sub1_df$look_stat,  X = PHI,
            nu=1.0, nlevel=1, NC.buffer = 2, findAwght = TRUE,
                          NC=20, LKinfo = LKinfo_stat)

LKinfo_stat$sigma
LKinfo_stat$rho
LKinfo_stat$lambda

sigma <- LKinfo_stat$sigma
rho   <- LKinfo_stat$rho
lambda<- LKinfo_stat$lambda

sigma^2/rho
```

