---
title: "Tensor_cov"
author: "Ryan Peterson"
date: "2024-03-17"
output: pdf_document
---

```{r library}
suppressMessages(library(LatticeKrig))
suppressMessages(library(viridis))
```

```{r test_data}
setwd("~/COS_LK")
load("DF_LK/DFSynthData_Base.rda")
```

Goal/Outline:

- Explore using tensor products for the basis function

-- Begin by exploring the basis function in LatticeKrig

--- Look into variance covariance plots (comparing reg and tensor)

- Place future work on tensors here. 


# LKrig.basis

## LKrig.cov

Looking into the behavior of the covariance.

```{r temp_function}
temp.fun <- function( NC.buffer=0, NC=4,  a.wght=4.01){
      LKinfo<- LKrigSetup(cbind( c(-1,1), c(-1,1)),nlevel=1, alpha=1,
                               a.wght=a.wght, NC=NC,   
                               NC.buffer=NC.buffer,
                              choleskyMemory=list(nnzR=2e6))
      cov1y<- cov1x<- cov0x<- cov0y<-  matrix( NA, nrow=200, ncol=20)
      cov1dx<- cov1dy<- cov0dx<- cov0dy<- matrix( NA, nrow=200, ncol=20)
      cgrid<- seq( 0,1,,20)
      for( k in 1:20){
          hold<- LKrig.cov.plot( LKinfo,
                          center=rbind( c(cgrid[k], cgrid[k])), NP=200)
          cov1x[,k] <- hold$cov[,1]
          cov1y[,k] <- hold$cov[,2]
          cov1dx[,k] <- hold$d[,1]
          cov1dy[,k] <- hold$d[,2]
          hold<- LKrig.cov.plot( LKinfo,
                           center=rbind( c(cgrid[k],0) ), NP=200)
          cov0x[,k] <- hold$cov[,1]
          cov0y[,k] <- hold$cov[,2]
          cov0dx[,k] <- hold$d[,1]
          cov0dy[,k] <- hold$d[,2]
              }
       matplot( cov1dx, cov1x, type="l", col= rainbow(20),
                       xlab="", ylab="correlation")
       mtext( side=1, line=-1, text="diagonal X")
       title( paste(  " buffer=",NC.buffer), cex=.5)
       matplot( cov1dy, cov1y, type="l", col= rainbow(20),
                      xlab="", ylab="")
       mtext( side=1, line=-1, text="diagonal Y")
       matplot(cov0dx, cov0x, type="l", col= rainbow(20),
                      xlab="",       ylab="")
       mtext( side=1, line=-1, text="middle X")
       matplot( cov0dy, cov0y, type="l", col= rainbow(20),
                       xlab="",   ylab="")
       mtext( side=1, line=-1, text="middle Y")
       title( paste( NC, a.wght), cex=.5)
}

temp.tensor.fun <- function( NC.buffer=0, NC=4,  a.wght=4.01){
      LKinfo<- LKrigSetup(cbind( c(-1,1), c(-1,1)),nlevel=1, alpha=1,
                               a.wght=a.wght, NC=NC,   
                               NC.buffer=NC.buffer,
                              choleskyMemory=list(nnzR=2e6),  BasisType="Tensor" )
      cov1y<- cov1x<- cov0x<- cov0y<-  matrix( NA, nrow=200, ncol=20)
      cov1dx<- cov1dy<- cov0dx<- cov0dy<- matrix( NA, nrow=200, ncol=20)
      cgrid<- seq( 0,1,,20)
      for( k in 1:20){
          hold<- LKrig.cov.plot( LKinfo,
                          center=rbind( c(cgrid[k], cgrid[k])), NP=200)
          cov1x[,k] <- hold$cov[,1]
          cov1y[,k] <- hold$cov[,2]
          cov1dx[,k] <- hold$d[,1]
          cov1dy[,k] <- hold$d[,2]
          hold<- LKrig.cov.plot( LKinfo,
                           center=rbind( c(cgrid[k],0) ), NP=200)
          cov0x[,k] <- hold$cov[,1]
          cov0y[,k] <- hold$cov[,2]
          cov0dx[,k] <- hold$d[,1]
          cov0dy[,k] <- hold$d[,2]
              }
       matplot( cov1dx, cov1x, type="l", col= rainbow(20),
                       xlab="", ylab="correlation")
       mtext( side=1, line=-1, text="diagonal X")
       title( paste(  " buffer=",NC.buffer), cex=.5)
       matplot( cov1dy, cov1y, type="l", col= rainbow(20),
                      xlab="", ylab="")
       mtext( side=1, line=-1, text="diagonal Y")
       matplot(cov0dx, cov0x, type="l", col= rainbow(20),
                      xlab="",       ylab="")
       mtext( side=1, line=-1, text="middle X")
       matplot( cov0dy, cov0y, type="l", col= rainbow(20),
                       xlab="",   ylab="")
       mtext( side=1, line=-1, text="middle Y")
       title( paste( NC, a.wght), cex=.5)
}
```



```{r LKrig.cov}
#TODO: tweak with the temp.functions to do comparison between radial/tensor
#running example from help file (LKrig.basis)
set.panel(3,4)
par(mar=c(3,4,1,0), oma=c(1,1,1,1))
temp.fun(  NC.buffer=5, NC=4, a.wght=4.05)
temp.fun(  NC.buffer=5, NC=16, a.wght=4.05)
temp.fun(  NC.buffer=5, NC=64, a.wght=4.05)

#repeat for tensor basis
set.panel(3,4)
par(mar=c(3,4,1,0), oma=c(1,1,1,1))
temp.tensor.fun(  NC.buffer=5, NC=4, a.wght=4.05)
temp.tensor.fun(  NC.buffer=5, NC=16, a.wght=4.05)
temp.tensor.fun(  NC.buffer=5, NC=64, a.wght=4.05)

##DO single run and compare numerically. 
```

```{r repeat_single}
#numeric check on tensor and radial covariance.
NC.buffer <- 0
NC <- 64
a.wght <- 4.01

LKinfo_radial <- LKrigSetup(cbind( c(-1,1), c(-1,1)), nlevel=1, alpha=1,
                               a.wght=a.wght, NC=NC,   
                               NC.buffer=NC.buffer,
                              choleskyMemory=list(nnzR=2e6))

##check to see if we need this
cov1y <- cov1x<-  matrix( NA, nrow=200, ncol=1)
#cov0x <- cov0y <-  matrix( NA, nrow=200, ncol=1)
cov1dx <- cov1dy<- matrix( NA, nrow=200, ncol=1)
#cov0dx<- cov0dy <- matrix( NA, nrow=200, ncol=1)
cgrid <- seq( 0,1,,20) #used as the based grid of center values choose from value 1-20

k = 1
hold<- LKrig.cov.plot( LKinfo_radial, center=rbind( c(cgrid[k], cgrid[k])), NP=200)
cov1x[,k] <- hold$cov[,1] #used in diagonal x
cov1y[,k] <- hold$cov[,2] #used in diagonal y
cov1dx[,k] <- hold$d[,1]  #used in diagonal x
cov1dy[,k] <- hold$d[,2]  #used in diagonal y


LKinfo_tensor<- LKrigSetup(cbind( c(-1,1), c(-1,1)),nlevel=1, alpha=1,
                               a.wght=a.wght, NC=NC,   
                               NC.buffer=NC.buffer,
                              choleskyMemory=list(nnzR=2e6),  BasisType="Tensor" )
cov1y_T<- cov1x_T<- matrix( NA, nrow=200, ncol=1)
cov1dx_T<- cov1dy_T<- matrix( NA, nrow=200, ncol=1)
cgrid<- seq( 0,1,,20)


hold_tensor<- LKrig.cov.plot( LKinfo_tensor, center=rbind( c(cgrid[k], cgrid[k])), NP=200)
cov1x_T[,k] <- hold_tensor$cov[,1]
cov1y_T[,k] <- hold_tensor$cov[,2]
cov1dx_T[,k] <- hold_tensor$d[,1]
cov1dy_T[,k] <- hold_tensor$d[,2]




cov1dx_new <- rbind(cov1dx, cov1dx_T)
cov1x_new <- rbind(cov1x, cov1x_T)

cov1dy_new <- rbind(cov1dy, cov1dy_T)
cov1y_new <- rbind(cov1y, cov1y_T)

matplot( cov1dx_new, cov1x_new, type="l", col= rainbow(20), 
         xlim = c(0.9, 1.0),  ylim = c(0.45,0.55),
                       xlab="", ylab="correlation")
matplot( cov1dy_new, cov1y_new, type="l", col= rainbow(20), 
         xlim = c(0.9, 1.0),  ylim = c(0.45,0.55),
                       xlab="", ylab="correlation")

#x_diff_4 <- cov1x_T - cov1x
#y_diff_4 <- cov1y_T - cov1y
set.panel(1,3)
plot(cov1dx, x_diff_4, type="l", col= rainbow(20), 
         xlim = c(0.0, 1.0), main = "NC = 4",
                       xlab="distance", ylab="correlation diff (tensor - radial)")


#x_diff_16 <- cov1x_T - cov1x
#y_diff_16 <- cov1y_T - cov1y
plot(cov1dx, x_diff_16, type="l", col= rainbow(20), 
         xlim = c(0.0, 1.0), main = "NC = 16",
                       xlab="distance", ylab="correlation diff (tensor - radial)")


x_diff_64 <- cov1x_T - cov1x
y_diff_64 <- cov1y_T - cov1y

plot(cov1dx, x_diff_64, type="l", col= rainbow(20), 
         xlim = c(0.0, 1.0), main = "NC = 64",
                       xlab="distance", ylab="correlation diff (tensor - radial)")

```

```{r}
#compare between pts around 0.70
cov1x_T[190:200] - cov1x[190:200]


```


## Basis Visualization

```{r basis_vis}
NC<- 10 
nlevel <- 4
a.wght <-  4 + 1/(.5)^2
alpha<-  1/2^(0:(nlevel-1)) 
LKinfo2<- LKrigSetup( cbind( c( -1,1), c(-1,1)), NC=NC,
                   nlevel=nlevel, a.wght=a.wght,alpha=alpha,
                   BasisType = "Radial")
LKinfo3 <- LKrigSetup( cbind( c( -1,1), c(-1,1)), NC=NC,
                   nlevel=nlevel, a.wght=a.wght,alpha=alpha,
                   BasisType="Tensor" )
# evaluate covariance  along the  horizontal line through
# midpoint of region -- (0,0) in this case. 
look<- LKrig.cov.plot( LKinfo2)
look2 <- LKrig.cov.plot( LKinfo3)
# a plot of the covariance function in x and y with respect to (0,0)
plot(look$u[,1], look$cov[,1], type="l", main = "X Transect")
lines(look2$u[,1], look2$cov[,1], type = "l", col = "darkmagenta")

plot(look$u[,2], look$cov[,2], type="l", main = "Y Transect")
lines(look2$u[,2], look2$cov[,2], type = "l", col = "darkmagenta")


#vis in 2-d (repeat for tensor)
set.panel(1,2)
x2<- cbind( 0,0)
x1<- make.surface.grid( list(x=seq( -1,1,,40),  y=seq( -1,1,,40)))
look<- LKrig.cov( x1,x2, LKinfo2)
contour( as.surface( x1, look), nlevels = 30)
title("Radial")

#2-d vis with tensor
x2<- cbind( 0,0)
x1<- make.surface.grid( list(x=seq( -1,1,,40),  y=seq( -1,1,,40)))
look2 <- LKrig.cov( x1,x2, LKinfo3)
contour( as.surface( x1, look2), nlevels = 30)
title("Tensor")
```


### Multiresolution Basis Functions

Adding in extra layers

```{r}
set.panel(3,2)
LKinfo<-LKrigSetup(cbind( c(-1,1), c(-1,1)), NC=6,
                    a.wght=5, alpha=c(1,.5,.25), nlevel=3,
                    NC.buffer=0)
# evaluate the basis functions on a grid to look at them
xtemp<- seq(-1,1,,40)
xg<- make.surface.grid( list(x=xtemp, y= xtemp) )
PHI<- LKrig.basis( xg,LKinfo)
# coerce to dense matrix format to make plotting easier.
PHI<- spam2full(PHI)
# first tenth, and last basis function in each resolution level
# basis functions centers are added
set.panel(3,3)
for(  j in 1:3){
  id1<- LKinfo$latticeInfo$offset[j]+ 1
  id2<-  LKinfo$latticeInfo$offset[j]+ 10
  idlast<- LKinfo$latticeInfo$offset[j] +
                  LKinfo$latticeInfo$mx[j,1]*LKinfo$latticeInfo$mx[j,2]
   
  centers<-  make.surface.grid(LKrigLatticeCenters(LKinfo, j) )
  image.plot( as.surface(xg, PHI[,id1]))
  points( centers, cex=.2, col="grey")
  image.plot(as.surface(xg, PHI[,id2]))
  points( centers, cex=.2, col="grey")
  image.plot( as.surface(xg, PHI[,idlast]))
  points( centers, cex=.2, col="grey")
}
  set.panel()
```


### Tensor Product Basis Function

```{r}
#LK setup
LKinfo<- LKrigSetup(cbind( c(-1,1), c(-1,1)), NC=6,
                                 nlevel=1, a.wght=4.5,alpha=1, NC.buffer=0 )
# evaluate the basis functions on a grid to look at them
xg<- make.surface.grid( list(x=seq(-1,1,,50), y= seq(-1,1,,50)))
PHI<- LKrig.basis( xg,LKinfo)
LKinfo2<- LKrigSetup(cbind( c(-1,1), c(-1,1)), NC=6,
                                 nlevel=1, a.wght=4.5,alpha=1, NC.buffer=0,
                                 BasisType="Tensor" )
PHI2<- LKrig.basis( xg,LKinfo2)



#adjust below visualization to be more centered (e.g. choose different basis function to look at)
image.plot(as.surface(xg, PHI[,16]))
points(  make.surface.grid( LKrigLatticeCenters(LKinfo, 1)) , col="grey", cex=.5, pch = 16)
title("A radial basis function")

image.plot(as.surface(xg, PHI2[,9]))
points(  make.surface.grid( LKrigLatticeCenters(LKinfo, 1)), col="grey", cex=.5, pch = 16)
title("Tensor product basis function")
  
image.plot(as.surface(xg, PHI2[,21] - PHI[,21]))
points(  make.surface.grid( LKrigLatticeCenters(LKinfo, 1)), col="grey", cex=.5)
title("Tensor - Radial (for 9th basis function)")     

#messing around with the sum of basis function
#TODO: expand this to undertsand overlap between tensor products (use PH2)
PHI_test <- PHI[,1] + PHI[,8] + PHI[,12]
image.plot(as.surface(xg, PHI_test))
```


# Testing COS with Tensor Product Basis Functions
Using testing methods from DF_Lk but only a single (simple) block data set.

```{r}

```

